### 2021-07-14 paper_draft/paper.tex
Added paper draft

\documentclass{article}
\usepackage[backend=biber, style=authoryear-icomp]{biblatex}
\usepackage{blindtext}
\usepackage{multicol}
\usepackage[document]{ragged2e}
\addbibresource{ref.bib}

\title{Using a Deep Autoencoder Model to Characterize Air Pollution}
\author{Nicolas Shannon$^{1}$, Bishnu Timalsena$^{2}$, Raul Chavez$^{3}$, \\Vandana Nunna Lakshmi$^{2}$}
\date{}
% To cite, use \parencite{} or \textcite{}

\begin{document}

\begin{centering}

\maketitle

\vspace{-0.50cm}
\scriptsize
{$^{1}$Department of Computer Science, Loyola University Chicago, Chicago, IL, US}\\
{$^{2}$Department of Computer Science, University North Texas, Denton, TX, US}\\
{$^{3}$Department of Electrical and Computer Engineering, University of Texas - Rio Grande Valley, Rio Grande Valley, TX, US}

\end{centering}
\begin{abstract}
Climate change is one of the most complex challenges humanity has ever faced. It is a multi-dimensional
problem that poses economic, social, scientific, political, and moral 
questions that must be addressed. One of the most important ways of 
combating climate change is to research pollution mitigation strategies 
so that policy makers and government planners are better equipped to  make 
informed decisions. We focused on the principle cause of global warming:
the polluting gases responsible for the rapid global climate shift.
We developed a technique to characterize cities by their levels of 
various polluting gases and particulate matter using an undercomplete autoencoder 
model. By creating a representation of pollution levels for numerous cities
across the United States, we explored some of the more prominent features 
that characterize urban pollution. We compared two techniques for reducing 
the dimensionality of our input - Principle Component Analysis (PCA) and
Undercomplete Autoencoders. The two first two hidden dimensions represented
a
large percentage of the explained variance in the model, so we clustered the cities in an effort to extract any useful features.
\end{abstract}

\vspace{0.25cm}

\section{Introduction}
\blindtext\blindtext

\vspace{0.25cm}

\begin{multicols}{2}
\section{Methodology}
The main use of an undercomplete autoencoder is not in its ability to perfectly reconstruct the input, but in the useful features it is capable of distilling.
In our case, we wanted to see what properties the autoencoder model could extract when given time series air pollution data for various cities throughout the United States. Using data from OpenWeather's Air Pollution API, we constructed data frames for eight polluting gases and particulates. We obtained the daily averages of each pollutant six and a half months for just over eighteen-thousand cities in the U.S. and other U.S. territories.
After obtaining the requisite data, we cleaned and preprocessed the data by removing cities that had the same name and any entries with missing values. While this made the data quickly useable, a more ideal solution would be to add a state/country code to distinguish between cities and to impute any missing values. For more on this see the future improvements section.
We compared the autoencoder model to Principle Component Analysis, a much older technique used for dimensionality reduction that has been a part of statistical literature since the early twentieth century.[2] The main distinction between autoencoders and PCA is that autoencoders can span a nonlinear subspace, whereas PCA learns the principle subspace.
Autoencoders that employ nonlinar encoder and decoder functions can create a more powerful generalization than PCA. However, an autoencoder model that is allowed too much capacity is at risk of reconstructing the input "too perfectly" without extracting any useful information.[3]
\end{multicols}

As \textcite{Goodfellow16} say, there's no reason...
\textcite{Jolliffe02}
\textcite{NAAQS09}

\printbibliography

\end{document}

### 2021-07-14 autoencoder.py

@@ -178,9 +178,9 @@ def grid_search(component, iter_dims, param_vec):
     # Write title and erase any previous values
     # Create title
     file_name = '/home/nicks/Documents/model_tuning_results/autoencoders/vector_dimensions/{}_vec_dim'.format(component)
-    with open(file_name,'w') as f:
-        writer = csv.writer(f)
-        writer.writerow(grid_list)
+    #with open(file_name,'w') as f:
+    #    writer = csv.writer(f)
+    #    writer.writerow(grid_list)
     # Perform a Grid Search (3875 cycles)
     for dim in iter_dims:
         best_var = -100
@@ -221,7 +221,7 @@ EPOCH = [10, 50, 75, 100]
 # Param vector
 PARAM_VEC = list(itertools.product(LR,BATCH,EPOCH))
 # List of key dimensions
-ITER_DIMS = np.concatenate((np.arange(2,26,1), np.array([30,40,50,60,80,100,120])), axis=0)
+ITER_DIMS = np.concatenate((np.arange(14,26,1), np.array([30,40,50,60,80,100,120])), axis=0)
 
 # Method tests
 #ae_run(DIMS, COMPONENT_NAMES)

### 2021-07-29 autoencoder.py
Added correlation matrix; fixed README; added way to save norm data in autoencoder

@@ -1,10 +1,8 @@
 import numpy as np
 import pandas as pd
 import itertools
-import matplotlib.pyplot as plt
 import csv
 
-from numpy.core.numeric import NaN
 from tensorflow.keras.models import Model
 from tensorflow.keras.layers import Dense, Input
 from tensorflow.keras.optimizers import Adam
@@ -40,14 +38,14 @@ def autoencoder(dfx, y, dim, component, activation, lr=0.0001, batch=128, epochs
     # Training and test splits
     train, dev, train_labels, dev_labels = train_test_split(dfx, y, test_size=0.20, random_state=40)
     train, test, train_labels, test_labels = train_test_split(train, train_labels, test_size=0.20, random_state=40)
-    input_data = Input(shape=(121,))
+    input_data = Input(shape=(191,))
 
     x_train = train.loc[:, train.columns]
     x_test = test.loc[:, test.columns]
     x_dev = dev.loc[:, dev.columns]
 
     encoded = Dense(dim, activation=activation[0], name='bottleneck')(input_data)
-    decoded = Dense(121, activation=activation[1])(encoded)
+    decoded = Dense(191, activation=activation[1])(encoded)
 
     autoencoder = Model(input_data, decoded)
 
@@ -102,18 +100,26 @@ def ae_run(dims, component_names):
         # Features list and removal of city, lat, lon
         features = list(df.columns.values)
         del features[:3] 
-        #del features[-1]
-        del features[121:]
-
+        del features[-1]
+        #del features[121:]
+        
         # y value list using last day of 7-month data
-        y = df.loc[:, ['{}_2021_03_28'.format(component)]].values
-        print(features)
+        y = df.loc[:, ['{}_2021_06_06'.format(component)]].values
         
         # Normalize x values; save in data frame
         x = df.loc[:, features].values
         x = Normalizer().fit_transform(x)
         dfx = pd.DataFrame(x)
 
+        ### Uncomment code below to save normalized data ###
+        # TODO: ^better solution for this; incorporate better
+        #norm_labels = ['dim_{}'.format(i) for i in range(1, dims+2)]
+        #dfx.columns = norm_labels
+        #
[…]

### 2021-07-29 visualization.py

@@ -1,7 +1,8 @@
 import pandas as pd
 import numpy as np
 import matplotlib.pyplot as plt
-
+import seaborn as sns
+import itertools
 
 
 def linegraph(f_name, type, dims, component_names, colors_list):
@@ -17,15 +18,15 @@ def linegraph(f_name, type, dims, component_names, colors_list):
     '''
     
     num_of_comp = list(range(2,dims+1))
+    # Plot results
     plt.figure(figsize=(12,10))
-    plt.rcParams.update({'font.size': 18})
     for i, component in enumerate(component_names):
         # AE or PCA
         if type == 'ae':
-            file_name = '%s/autoencoders/trial2/%s_ae_results.csv' % (f_name, component)
+            file_name = '{FILE}{COMPONENT}_ae_results_gs.csv'.format(FILE=f_name, COMPONENT=component)
             plt_title = 'Autoencoder Reduced Representation of Air Pollutants'
         elif type == 'pca':
-            file_name = '%s/autoencoders/trial2/%s_pca_results.csv' % (f_name, component)
+            file_name = '{FILE}{COMPONENT}_pca_results_gs.csv'.format(FILE=f_name, COMPONENT=component)
             plt_title = 'PCA Reduced Representation of Air Pollutants'
         else:
             print('Type must be "ae" or "pca"')
@@ -36,12 +37,17 @@ def linegraph(f_name, type, dims, component_names, colors_list):
         variance = model['{}_var'.format(component)]
         r2 = model['{}_r2'.format(component)]
         
-        # Plot results
+
         plt.plot(num_of_comp, variance[:dims-1], label = '{}'.format(component), linestyle = '-', marker = '+', color = colors_list[i])
         plt.plot(num_of_comp, r2[:dims-1], linestyle = '-.', marker = 'H', color = colors_list[i])
-        plt.xlabel('Dimension')
-        plt.ylabel('% Explained Variance')
-        plt.title(plt_title)
+        plt.rcParams.update({'font.size': 22})
+        plt.tick_params(axis='both', which='major', labelsize=28)
+       # plt.xlabel('Dimension')
+       # plt.ylabel('% Explained Variance')
+        plt.xlabel('')
+        plt.ylabel('')
[…]

### 2021-07-29 paper-draft/paper.tex
Added backup of paper

\documentclass{article}
\usepackage[backend=biber, style=authoryear-icomp]{biblatex}
\usepackage{blindtext}
\usepackage{multicol}
\usepackage[document]{ragged2e}
\usepackage[pass, letterpaper]{geometry}
%\usepackage[margin=1in]{geometry}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{subcaption}
\addbibresource{ref.bib}

% Define macros
\newcommand\note[1]{\textbf{\textcolor{red}{#1}}}

% Captions
\captionsetup{justification = raggedright, singlelinecheck = false, labelfont=bf}

\title{Using a Deep Autoencoder Model to Characterize Air Pollution}
\author{Nicolas Shannon$^{1}$, Bishnu Timalsena$^{2}$, Raul Chavez Jr.$^{3}$, \\Vandana Nunna Lakshmi$^{2}$}
\date{}
% To cite, use \parencite{} or \textcite{}

\begin{document}

% Acknowledgements
\begin{centering}
\maketitle
\vspace{-0.50cm}
\scriptsize
{$^{1}$Department of Computer Science, Loyola University Chicago, Chicago, IL, US}\\
{$^{2}$Department of Computer Science, University North Texas, Denton, TX, US}\\
{$^{3}$Department of Electrical and Computer Engineering, University of Texas 
    - Rio Grande Valley, Brownsville, TX, US \note{$\leftarrow$ fix formatting}}
\end{centering}

\begin{abstract}
Air pollutants are both a cause and a symptom of climate change. Global warming has the potential to raise
the levels of various air pollutants. The rise in surface air pollutants can lead to heat trapped close to 
the Earth's surface by virtue of the greenhouse effect, which further perpetuates the cycle. In order to
more effectively combat climate change, it is important to understand and characterize the source of 
emissions. We developed a technique to characterize cities by the levels of various air toxins using an
undercomplete autoencoder model. By representing the pollution levels of eighteen thousand cities across
the United States, we explore some of the more prominent features that characterize urban pollution.
We compare two techniques for reducing the input dimensionality - Principle Component Analysis (PCA) and
Undercomplete Autoencoders.
\end{abstract}

\vspace{0.25cm}
\section{Introduction}
Over the past several decades, air pollution has been a major concern for human health
and the environment. The EPA sets national air quality standards for  six major air
toxins: ground level Ozone (O$_3$), Carbon Monoxide (CO), Sulfur Dioxide (SO$_2$),
Lead (Pb), Nitrogen Dioxide (NO$_2$), and particulate matter (PM). \parencite{NAAQS09}.
The sources of these emissions vary, but the two largest contributors are industrial 
processes and motorized vehicles.

\newpage

Short and long term exposure to air toxins pose significant risk to human health. 
A study by \textcite{Wang19} found a correlation  between increased ambient CO
levels and outpatient visits for respiratory, cardiovascular, genitourinary, and gastrointenstinal
diseases. Several studies performed by the EPA also suggest that people with pre-existing health
conditions are at higher risk when exposed to CO over long periods.\par

\subsection{Air Toxins}
We wanted to explore how CO and other gases can be characterized with respect to location
in an effort to provide researchers and policy makers with the context and tools necessary to
make informed decisions. There are a litany of technologies and tools for pollution prevention
and control, but most of these options require contextual information and large amounts of data.
\par
By creating an embedding of time series pollution data for numerous cities throughout the U.S.,
we can explore the more interesting features of the model as well as apply it to tangential 
areas such as the pollution generated by industrial farming and animal husbandry activities.

\begin{figure}[h]
\includegraphics[width=\textwidth]{/home/nicks/Documents/paper-draft/images/co-concentration-table.png}
    \caption{Carbon Monoxide Exposure Table, Parts Per Million (ppm)}
    \label{fig:co-exposure}
\end{figure}

\subsection{A Deep Learning Approach}
In this paper we compare the autoencoder model with principle component analysis (PCA), a much
older technique for dimensional reduction. Using a deep autoencoder has several advantages.
First, deep autoencoders are better at compressing highly complex data into few dimensions.
Autoencoders with a nonlinear encoder and decoder function can learn more powerful nonlinear
generalizations than PCA. \parencite{Goodfellow16}.
\par \note{Autoencoder grapic from poster here}

\section{Related Work}
The idea of generating embeddings for a location is a highly potential expanse for research.
The two major forces behind its competency are, the fact that large volumes of data is being
effectively compressed and that these embeddings, when plugged into any model could effectively
represent any given location.  A lot of research is being done on location embeddings in recent
times. For instance, location embeddings are being generated based on human-mobility data 
(Crivellari, A.; 2019) to understand the behavioral proximity between different locations 
irrespective of their spatial distances. The sequence of locations that are frequently
travelled were fed to a Skip-gram Word2vec model to generate the location embeddings. Another 
approach that was implemented to leverage location embeddings is, GPS2Vec ( Yifang Yin, 2019).
This application consists of a neural network that would generate embeddings for a location, 
based on the semantic contexts that are sourced from the user content published on multiple 
digital platforms. With the integration of these embeddings, a successful geo-tagged image
classification task is demonstrated. In another effort (Carl Yang, 2019), the location 
embeddings derived from user content such as addresses, phone numbers, images, place names and 
all other details from multiple individual online posts, are compressed to form low dimensional
embeddings to solve the place duplication issue. Deduplication of locations, is a paramount 
task for any digital platform since they provide a place graph to its users, which is the 
result of merging location data from multiple sources. This merge leads to duplicity since 
different sources could hold different names for the same place. By applying K-NN search, 
pair-wise duplication prediction on these generated location embeddings, the application can
identify the duplicates and eliminate them to result in an efficient place graph.

\section{Methodology}

We used two different methods to reduce the dimensionality of our data - Principle Component
Analysis and Undercomplete Autoencoders. A simple linear regression was performed on 
the encoded data that was generated by both the PCA and AE models. Additionally, we performed
a linear regression on the unencoded values for comparison. Each model's goodness of fit
was evaluated using the percentage of explained variance for the three linear regressions.

% pipeline
\begin{figure}[h]
\includegraphics[width=\textwidth, height=6cm]{/home/nicks/Documents/paper-draft/images/pipeline.png}
\caption{Data Pipeline}
\label{fig:pipeline}
\end{figure}

\subsection{Data Collection and Preprocessing}
The dataset was created using OpenWeather's Air Pollution API. The dataset includes data for 
eight polluting gases and particulates: CO, NO, NO$_2$, O$_3$, SO$_2$, NH$_3$, PM$_{2.5}$, 
and PM$_{10}$. The dataset also includes the hourly timestamp, the latitude and longitude (part of the input),
and the air quality index, the latter of which we did not end up using.\par
We initially looked at data for the eight pollutants over a six-and-a-half month time 
frame: January 27th, 2020 through July 6th, 2021. However, we found that we achieved
better results with just four months of of data for our input features. Our final 
dataset included just over eighteen thousand cities in the United States and other U.S. 
territories. Additionally, we found it easier to focus on a single gas when developing 
the models, so we chose to concentrate on CO during the engineering process. Given
enough time and computational power, the techniques performed on the CO model would be 
easily extensible and transferable to the other seven pollutants.

% dataset characteristics table
\begin{table}[h!]
    \caption{Dataset Characteristics}
    \label{tab:table1}
    \vspace{0.1cm}
    \begin{tabular}{p{4cm}p{7cm}}
        \hline
        \multicolumn{2}{c}{OpenWeather Pollution API} \\
        \hline
        Time period  & 6.5 months    \\
        Time frequency & Daily  \\
        $\#$ of cities & 18,526 \\
        $\#$ of features & 120 (originally 190) \\
        $\#$ of component gases & 8 \\
        Dependent variable & Single day of time series (3/27/21) \\
        \hline
    \end{tabular}
\end{table}

OpenWeather's Pollution API ordinarily returns hourly values, but we opted to use the daily
average pollution levels as our feature set. We resolved to feature engineer the daily 
averages and to reduce the time scale to four months in order to keep the data 
proportional to the number of features and thus avoid the curse of dimensionality. 
\parencite{Trunk79}. \note{Idk about this} \par
We separated the dependent from the input before unit normalizing the input. 
The normalized data was then encoded by each model.

\subsection{Autoencoder and PCA Models}
The autoencoder model used consists of an encoded and decoded layer. We performed
a grid search for the CO model to determine the optimal hyperparameters. The following 
hyperparameters were tested:
\begin{itemize}
    \item \textit{Learning Rate:} 0.0001, 0.001, 0.01, 0.1
    \item \textit{Batch Size:} 32, 64, 128, 256
    \item \textit{Epochs:} 10, 50, 75, 100
\end{itemize}
We used the parameters that appeared most frequently over the one hundred twenty dimensions and 
applied them over the entire range so as not to overfit.\note{$\leftarrow$ not sure if this is correct}

\begin{table}[h!]
    \caption{Dataset Characteristics}
    \label{tab:table1}
    \vspace{0.1cm}
    \begin{tabular}{p{4cm}p{7cm}}
        \hline
        \multicolumn{2}{c}{Optimal Hyperparameters for Carbon Monoxide (CO)} \\
        \hline
        Activation function & Leaky ReLU \\
        Loss function & Mean Squared Error  \\
        Optimizer & Adam \\
        Learning rate & 0.001 \\
        Batch size & 64 \\
        Epochs & 150 \\ 
        \hline
    \end{tabular}
\end{table}

The PCA model was trained with k-fold cross-validation using five folds. The dataset was
split into a training, validation, and hold-out test set.

\section{Results}
First, we wanted to explore the correlation between days in our dataset. We calculated
the correlation coefficients and created a correlation matrix of the days that had
correlation coefficients of the highest magnitude.\par

% pipeline
\begin{figure}[h]
\includegraphics[width=\textwidth]{/home/nicks/Documents/paper-draft/images/corr_matrix_25.png}
\caption{Correlation Matrix}
\label{fig:pipeline}
\end{figure}

We compared the percent explained variance across one hundred twenty dimensions for
the encoded values generated by the PCA and autoencoder models as well as for the unencoded
values by using a simple linear regression. While PCA was overall less noisy, we found that
the autoencoder model achieved a much higher explainable variance for the first few dimensions.
\\ \note{Note that results graphs are placeholders}
% ae vs pca linegraph
\begin{figure}[h!]

\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nicks/Documents/paper-draft/images/ae_all_dims.png} 
    \caption{Autoencoder Dimensionality Reduction}
    \label{fig:ae_dim_reduction}
\end{subfigure}% <- % sign must be here!
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nicks/Documents/paper-draft/images/ae_all_dims.png}
    \caption{PCA Dimensionality Reduction}
    \label{fig:pca_dim_reduction}
\end{subfigure}

\caption{Graph of autoencoder and PCA reduced representations of air pollutants}
\label{fig:ae_vs_pca}
\end{figure}

\par We clustered the cities based on the first two latent dimensions in order to see how
the model characterizes each city

% co scatter
\begin{figure}[h!]

\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nicks/Documents/paper-draft/images/co_scatter_ae_wlist.png} 
    \caption{Outlier Cities}
    \label{fig:outliers}
\end{subfigure}% <- % sign must be here!
\begin{subfigure}{0.5\textwidth}
    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nicks/Documents/paper-draft/images/co_scatter_ae_large_cities.png}
    \caption{Densely Populated Cities}
    \label{fig:dense_cities}
\end{subfigure}

\caption{Scatter plots of outlier and highly-populated cities over first two dimensions}
\label{fig:outliers_vs_dense_cities}
\end{figure}

By selecting from a list the top two hundred highly-populated cities, we found that nearly
all of them were located near the center of the clustering. Conversely, we found that many 
of the outliers were smaller cities in rural locations.

\section{Discussion}

\section{Conclusion}

\printbibliography

\end{document}

### 2021-08-04 preprocessing.py
Improved functionality and readability of preprocessing, added a .gitignore

@@ -1,24 +1,22 @@
 import requests
 import json
 import pandas as pd
-import config
-import matplotlib.pyplot as plt
 import datetime
 import asyncio
+import os
+from dotenv import load_dotenv
 from csv import writer
 
-# Version 2 of preprocessing
-# generalizes and obtains lists for each particulate/gas component
-
+load_dotenv()   # Load environment variables
 
 def gen_daily_data(name, lat, lon, t_start, t_end, component_names):
     '''
     Generate data for a given gas/particulate for a given city over a set time period and write to csv
 
     @params: 
-        name: name of location
-        lat, lon: latitude and longitude in decimal degrees
-        t_start, t_end: starting and ending epoch in Unix time
+        name: Name of location
+        lat, lon: Latitude and longitude in decimal degrees
+        t_start, t_end: Starting and ending epoch in Unix time
     '''
     
     # Connect to endpoint and load data
@@ -27,7 +25,7 @@ def gen_daily_data(name, lat, lon, t_start, t_end, component_names):
         LON=lon,
         START=t_start,
         END=t_end,
-        KEY=config.OPEN_WEATHER_KEY
+        KEY=os.getenv('OPEN_WEATHER_KEY')
     )
     page = requests.get(url=endpoint)
     content = json.loads(page.content)
@@ -64,13 +62,13 @@ def gen_cols(t_start, t_end, component):
     Derive number of entries from start and end
 
     @params:
-        t_start, t_end: start and end times in unix time
-        element: string to represent the gas/particulate name to use
+        t_start, t_end: Start and end times in unix time
+        element: String to represent the gas/particulate name to use
     '''
 
     # Change in epoch to number of hours gets us total entries
     num_entries = int((t_end - t_start) / 86400)
-    time_step = T_START
+    time_step = t_start
     col_names = ['city', 'lat', 'lon']
 
     for i in range(num_entries):
@@ -85,20 +83,21 @@ def gen_cols(t_start, t_end, component):
     return col_names
 
 
-def batch_request(city_df, component_names):
+def batch_request(f_name, t_start, t_end, city_df, component_names):
     ''' 
     Places a batch request for gas/particulate levels on every minute 
 
     @params:
-        col_names: list of columns including city_name, lat/lon, and PM2.5 features
-        city_df: data frame containing city information
+        f_name: Name of files to write to
[…]

### 2021-08-04 autoencoder.py
Updated .gitignore, moved project files to src directory, cleaned up preprocessing

@@ -1,235 +1,0 @@
-import numpy as np
-import pandas as pd
-import itertools
-import csv
-
-from tensorflow.keras.models import Model
-from tensorflow.keras.layers import Dense, Input
-from tensorflow.keras.optimizers import Adam
-from sklearn.preprocessing import Normalizer
-from sklearn.linear_model import LinearRegression
-from sklearn.model_selection import train_test_split
-from sklearn.metrics import r2_score
-
-
-def autoencoder(dfx, y, dim, component, activation, lr=0.0001, batch=128, epochs=50):
-    '''
-    Autoencoder model
-    
-    @params:
-        dfx: DataFrame of normalized x values
-        y: Dependent variable to be predicted
-        dim: Current dimension
-        component: Name of gas or particulate
-        activation: Activation function represented as a tuple pair
-        lr: Learning rate, default of 0.0001
-        batch: Batch size, default of 128
-        epochs: Number of epochs, default of 50
-    @return:
-        variance_list: List of explainable variance scores
-        r2_list: List of R squared scores
-        encoded_data: The encoded X training data
-    '''
-
-    # Define the linear regression model
-    regr = LinearRegression()
-
-    print('---------- Autoencoder dim {} for {} ----------'.format(dim, component))
-    # Training and test splits
-    train, dev, train_labels, dev_labels = train_test_split(dfx, y, test_size=0.20, random_state=40)
-    train, test, train_labels, test_labels = train_test_split(train, train_labels, test_size=0.20, random_state=40)
-    input_data = Input(shape=(191,))
-
-    x_train = train.loc[:, train.columns]
-    x_test = test.loc[:, test.columns]
-    x_dev = dev.loc[:, dev.columns]
-
-    encoded = Dense(dim, activation=activation[0], name='bottleneck')(input_data)
-    decoded = Dense(191, activation=activation[1])(encoded)
-
-    autoencoder = Model(input_data, decoded)
-
-    # Compile the autoencoder
-    autoencoder.compile(loss='mean_squared_error', optimizer=Adam(learning_rate=lr))
-    autoencoder.summary()
-    
-    trained_model = autoencoder.fit(
-        x_train, 
-        x_train, 
-        batch_size=batch, 
-        epochs=epochs, 
-        verbose=1,
-        validation_data=(x_dev, x_dev)
-    )
-    
-    # Bottleneck representation
[…]

### 2021-08-04 config.py

@@ -1,2 +1,0 @@
-# API Keys
-OPEN_WEATHER_KEY = 'be8a45eab725130570596e89e88f7be7'

### 2021-08-04 linreg.py

@@ -1,94 +1,0 @@
-import numpy as np
-from numpy.core.numeric import NaN
-import pandas as pd
-import matplotlib.pyplot as plt
-
-from tensorflow.keras.models import Model
-from tensorflow.keras.layers import Dense, Input
-from tensorflow.keras.optimizers import Adam
-from sklearn.preprocessing import Normalizer
-from sklearn.linear_model import LinearRegression
-from sklearn.model_selection import train_test_split
-from sklearn.metrics import r2_score
-
-
-def linreg(dfx, y, component):
-    '''
-    Simple linear regression on uncompressed values
-    
-    @params:
-        dfx: DataFrame of normalized x values
-        y: Predicted value
-        component: Name of gas or particulate
-    '''
-
-    regr = LinearRegression()
-    variance_list = []
-    r2_list = []
-    
-    # Training and test splits
-    train, dev, train_labels, dev_labels = train_test_split(dfx, y, test_size=0.20, random_state=40)
-    train, test, train_labels, test_labels = train_test_split(train, train_labels, test_size=0.20, random_state=40)
-    input_data = Input(shape=(191,))
-
-    X_train = train.loc[:, train.columns]
-    X_test = test.loc[:, test.columns]
-    X_dev = dev.loc[:, dev.columns]
-
-    # Variance score explanation
-    regr.fit(train, train_labels)
-
-    Y_pred = regr.predict(X_test) # << Pass x_test to get predicitons for original uncompressed
-
-    # Variance and r2 scores for the regression model
-    variance = regr.score(X_test, test_labels) # << pass in x test for uncompressed values
-    r2_val = r2_score(test_labels, Y_pred)
-    print ('Variance score: %.2f' % variance)
-    print ('R Square', r2_val)
-
-    # Add r2/variance to lists
-    variance_list.append(variance)
-    r2_list.append(r2_val)        
-
-    print('variance:',variance_list)
-    print('r2: ', r2_list)
-    print('x test: ', X_test)
-    return (variance_list, r2_list, X_test)
-
-
-def linreg_run(component_names):
-    '''
-    Run the linear regression for each component gas
-
-    @params:
-        component_names: Gas/particulate name list
-    '''
-    for component in component_names:
-        print('---------- Running Linear Regression for {} ----------'.format(component))
[…]

### 2021-08-04 pca.py

@@ -1,209 +1,0 @@
-import pandas as pd
-import numpy as np
-import matplotlib.pyplot as plt
-
-from sklearn.decomposition import PCA
-from sklearn.model_selection import KFold, train_test_split
-from sklearn.metrics import r2_score
-from sklearn.linear_model import LinearRegression
-from sklearn.preprocessing import Normalizer
-from sklearn.cluster import KMeans
-
-
-def pca(i, X_train, X_test, Y_train, Y_test, folds=0, dev=False):
-    '''
-    Perform Linear Regression on either train/validation set or on test set for PCA
-
-    @params:
-        i: number of components
-        X_train, X_test, Y_train, Y_test: testing/training lists
-        dev: set to true if doing validation testing; false if using test set data      
-    '''
-
-    # Perform PCA on data set
-    pca = PCA(n_components=i)
-    X_train = pca.fit_transform(X_train)
-    X_test = pca.transform(X_test)
-    model = LinearRegression()
-
-    # Linear regression on validation/train set
-    if dev == True:
-        if i%10 == 0:
-            print('PCA validation -- fold {} -- component {}'.format(folds,i))
-        train_pca, dev_pca, train_labels_pca, dev_labels_pca = train_test_split(X_train, Y_train, test_size=0.33, random_state=42)
-        model.fit(train_pca, train_labels_pca)
-        res_sum_square= np.mean((model.predict(dev_pca) - dev_labels_pca ** 2))
-        var_val=model.score(dev_pca, dev_labels_pca)
-        Y_test_predict = model.predict(dev_pca)
-        r2_val=r2_score(dev_labels_pca,Y_test_predict)
-        return (model.intercept_, model.coef_, res_sum_square, var_val, r2_val)
-
-    # Linear regression on test set
-    else:
-        if i%10 == 0:
-            print('PCA Test -- component {}'.format(i))
-        model.fit(X_train, Y_train)
-        res_sum_square= np.mean((model.predict(X_test) - Y_test ** 2))
-        var_val=model.score(X_test, Y_test)
-        Y_test_predict = model.predict(X_test)
-        r2_val=r2_score(Y_test,Y_test_predict)
-        return (var_val, r2_val, X_train)
-
-
-def pca_run(dims, component_names):
-    ''' 
-    Run PCA using k-fold cross validation strategy
-
-    @params:
-        dims: Number of starting dimensions
-        component_names: The names of the gases/particulates
-    '''
-
-    for component in component_names:
[…]

### 2021-08-04 preprocessing.py

@@ -1,196 +1,0 @@
-import requests
-import json
-import pandas as pd
-import datetime
-import asyncio
-import os
-from dotenv import load_dotenv
-from csv import writer
-
-load_dotenv()   # Load environment variables
-
-def gen_daily_data(name, lat, lon, t_start, t_end, component_names):
-    '''
-    Generate data for a given gas/particulate for a given city over a set time period and write to csv
-
-    @params: 
-        name: Name of location
-        lat, lon: Latitude and longitude in decimal degrees
-        t_start, t_end: Starting and ending epoch in Unix time
-    '''
-    
-    # Connect to endpoint and load data
-    endpoint = 'http://api.openweathermap.org/data/2.5/air_pollution/history?lat={LAT}&lon={LON}&start={START}&end={END}&appid={KEY}'.format(
-        LAT=lat, 
-        LON=lon,
-        START=t_start,
-        END=t_end,
-        KEY=os.getenv('OPEN_WEATHER_KEY')
-    )
-    page = requests.get(url=endpoint)
-    content = json.loads(page.content)
-    df = pd.json_normalize(content)
-    
-    # List all records
-    ls = df['list'][0]
-    df_size = len(ls)
-    
-    # Hourly time series list of each entry; daily total of each entry
-    ts_list = []
-    component_dict = {}
-    total = 0
-
-    # Find the daily averages of each component; merge into component list
-    for component in component_names:
-        for i in range(df_size):
-            total+=ls[i]['components'][component]
-            # Average for each day
-            if (i%24 == 0):   
-                daily_value = round(total/24, 5)
-                ts_list.append(daily_value)
-                total = 0
-        # Each time series list goes into a new dict entry
-        component_dict[component] = ts_list
-        # Clear list for next iteration
-        ts_list = []
-    return component_dict
-
-
-def gen_cols(t_start, t_end, component):
-    ''' 
-    Get list of column names based on the number of entries (each daily average of data will be one column)
-    Derive number of entries from start and end
-
-    @params:
-        t_start, t_end: Start and end times in unix time
-        element: String to represent the gas/particulate name to use
-    '''
-
-    # Change in epoch to number of hours gets us total entries
-    num_entries = int((t_end - t_start) / 86400)
-    time_step = t_start
-    col_names = ['city', 'lat', 'lon']
-
-    for i in range(num_entries):
-        # Convert daily interval to human-readable
[…]

### 2021-08-04 src/autoencoder.py

@@ -1,0 +1,227 @@
+import numpy as np
+import pandas as pd
+import itertools
+import csv
+
+from tensorflow.keras.models import Model
+from tensorflow.keras.layers import Dense, Input
+from tensorflow.keras.optimizers import Adam
+from sklearn.preprocessing import Normalizer
+from sklearn.linear_model import LinearRegression
+from sklearn.model_selection import train_test_split
+from sklearn.metrics import r2_score
+
+
+def autoencoder(dfx, y, dim, component, activation, lr=0.0001, batch=128, epochs=50):
+    '''
+    Autoencoder model
+    
+    @params:
+        dfx: DataFrame of normalized x values
+        y: Dependent variable to be predicted
+        dim: Current dimension
+        component: Name of gas or particulate
+        activation: Activation function represented as a tuple pair
+        lr: Learning rate, default of 0.0001
+        batch: Batch size, default of 128
+        epochs: Number of epochs, default of 50
+    @return:
+        variance_list: List of explainable variance scores
+        r2_list: List of R squared scores
+        encoded_data: The encoded X training data
+    '''
+
+    # Define the linear regression model
+    regr = LinearRegression()
+
+    print('---------- Autoencoder dim {} for {} ----------'.format(dim, component))
+    # Training and test splits
+    train, dev, train_labels, dev_labels = train_test_split(dfx, y, test_size=0.20, random_state=40)
+    train, test, train_labels, test_labels = train_test_split(train, train_labels, test_size=0.20, random_state=40)
+    input_data = Input(shape=(191,))
+
+    x_train = train.loc[:, train.columns]
+    x_test = test.loc[:, test.columns]
+    x_dev = dev.loc[:, dev.columns]
+
+    encoded = Dense(dim, activation=activation[0], name='bottleneck')(input_data)
+    decoded = Dense(191, activation=activation[1])(encoded)
+
+    autoencoder = Model(input_data, decoded)
+
+    # Compile the autoencoder
+    autoencoder.compile(loss='mean_squared_error', optimizer=Adam(learning_rate=lr))
+    autoencoder.summary()
+    
+    trained_model = autoencoder.fit(
+        x_train, 
+        x_train, 
+        batch_size=batch, 
+        epochs=epochs, 
+        verbose=1,
+        validation_data=(x_dev, x_dev)
+    )
+    
+    # Bottleneck representation
[…]

### 2021-08-04 src/linreg.py

@@ -1,0 +1,94 @@
+import numpy as np
+from numpy.core.numeric import NaN
+import pandas as pd
+import matplotlib.pyplot as plt
+
+from tensorflow.keras.models import Model
+from tensorflow.keras.layers import Dense, Input
+from tensorflow.keras.optimizers import Adam
+from sklearn.preprocessing import Normalizer
+from sklearn.linear_model import LinearRegression
+from sklearn.model_selection import train_test_split
+from sklearn.metrics import r2_score
+
+
+def linreg(dfx, y, component):
+    '''
+    Simple linear regression on uncompressed values
+    
+    @params:
+        dfx: DataFrame of normalized x values
+        y: Predicted value
+        component: Name of gas or particulate
+    '''
+
+    regr = LinearRegression()
+    variance_list = []
+    r2_list = []
+    
+    # Training and test splits
+    train, dev, train_labels, dev_labels = train_test_split(dfx, y, test_size=0.20, random_state=40)
+    train, test, train_labels, test_labels = train_test_split(train, train_labels, test_size=0.20, random_state=40)
+    input_data = Input(shape=(191,))
+
+    X_train = train.loc[:, train.columns]
+    X_test = test.loc[:, test.columns]
+    X_dev = dev.loc[:, dev.columns]
+    
+    # Variance score explanation
+    regr.fit(train, train_labels)
+
+    Y_pred = regr.predict(X_test) # << Pass x_test to get predicitons for original uncompressed
+
+    # Variance and r2 scores for the regression model
+    variance = regr.score(X_test, test_labels) # << pass in x test for uncompressed values
+    r2_val = r2_score(test_labels, Y_pred)
+    print ('Variance score: %.2f' % variance)
+    print ('R Square', r2_val)
+
+    # Add r2/variance to lists
+    variance_list.append(variance)
+    r2_list.append(r2_val)        
+
+    print('variance:',variance_list)
+    print('r2: ', r2_list)
+    print('x test: ', X_test)
+    return (variance_list, r2_list, X_test)
+
+
+def linreg_run(component_names):
+    '''
+    Run the linear regression for each component gas
+
+    @params:
+        component_names: Gas/particulate name list
+    '''
+    for component in component_names:
+        print('---------- Running Linear Regression for {} ----------'.format(component))
[…]

### 2021-08-04 src/pca.py

@@ -1,0 +1,209 @@
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+
+from sklearn.decomposition import PCA
+from sklearn.model_selection import KFold, train_test_split
+from sklearn.metrics import r2_score
+from sklearn.linear_model import LinearRegression
+from sklearn.preprocessing import Normalizer
+from sklearn.cluster import KMeans
+
+
+def pca(i, X_train, X_test, Y_train, Y_test, folds=0, dev=False):
+    '''
+    Perform Linear Regression on either train/validation set or on test set for PCA
+
+    @params:
+        i: number of components
+        X_train, X_test, Y_train, Y_test: testing/training lists
+        dev: set to true if doing validation testing; false if using test set data      
+    '''
+
+    # Perform PCA on data set
+    pca = PCA(n_components=i)
+    X_train = pca.fit_transform(X_train)
+    X_test = pca.transform(X_test)
+    model = LinearRegression()
+
+    # Linear regression on validation/train set
+    if dev == True:
+        if i%10 == 0:
+            print('PCA validation -- fold {} -- component {}'.format(folds,i))
+        train_pca, dev_pca, train_labels_pca, dev_labels_pca = train_test_split(X_train, Y_train, test_size=0.33, random_state=42)
+        model.fit(train_pca, train_labels_pca)
+        res_sum_square= np.mean((model.predict(dev_pca) - dev_labels_pca ** 2))
+        var_val=model.score(dev_pca, dev_labels_pca)
+        Y_test_predict = model.predict(dev_pca)
+        r2_val=r2_score(dev_labels_pca,Y_test_predict)
+        return (model.intercept_, model.coef_, res_sum_square, var_val, r2_val)
+
+    # Linear regression on test set
+    else:
+        if i%10 == 0:
+            print('PCA Test -- component {}'.format(i))
+        model.fit(X_train, Y_train)
+        res_sum_square= np.mean((model.predict(X_test) - Y_test ** 2))
+        var_val=model.score(X_test, Y_test)
+        Y_test_predict = model.predict(X_test)
+        r2_val=r2_score(Y_test,Y_test_predict)
+        return (var_val, r2_val, X_train)
+
+
+def pca_run(dims, component_names):
+    ''' 
+    Run PCA using k-fold cross validation strategy
+
+    @params:
+        dims: Number of starting dimensions
+        component_names: The names of the gases/particulates
+    '''
+
+    for component in component_names:
[…]

### 2021-08-04 src/preprocessing.py

@@ -1,0 +1,245 @@
+import requests
+import json
+import pandas as pd
+import datetime
+import asyncio
+import os
+from sklearn.preprocessing import Normalizer
+import logging
+import sys
+from dotenv import load_dotenv
+from csv import writer
+
+
+load_dotenv()   # Load environment variables
+
+def gen_daily_data(name, lat, lon, t_start, t_end, component_names):
+    '''
+    Generate data for a given gas/particulate for a given city over a set time period and write to csv
+
+    @params: 
+        name: Name of location
+        lat, lon: Latitude and longitude in decimal degrees
+        t_start, t_end: Starting and ending epoch in Unix time
+    '''
+    
+    # Connect to endpoint and load data
+    endpoint = 'http://api.openweathermap.org/data/2.5/air_pollution/history?lat={LAT}&lon={LON}&start={START}&end={END}&appid={KEY}'.format(
+        LAT=lat, 
+        LON=lon,
+        START=t_start,
+        END=t_end,
+        KEY=os.getenv('OPEN_WEATHER_KEY')
+    )
+    page = requests.get(url=endpoint)
+    content = json.loads(page.content)
+    df = pd.json_normalize(content)
+    
+    # List all records
+    ls = df['list'][0]
+    df_size = len(ls)
+    
+    # Hourly time series list of each entry; daily total of each entry
+    ts_list = []
+    component_dict = {}
+    total = 0
+
+    # Find the daily averages of each component; merge into component list
+    for component in component_names:
+        for i in range(df_size):
+            total+=ls[i]['components'][component]
+            # Average for each day
+            if (i%24 == 0):   
+                daily_value = round(total/24, 5)
+                ts_list.append(daily_value)
+                total = 0
+        # Each time series list goes into a new dict entry
+        component_dict[component] = ts_list
+        # Clear list for next iteration
+        ts_list = []
+    return component_dict
+
+
+def unit_norm(file_in, file_out, component_names, dims):
+    '''
+    Unit Normalize the gas data on a 0-1 scale and save to file
+
+    @params:
+        file_in: File containing input data
+        file_out: Destination to output unit normalized data
+        component_names: List of component gases
+        dims: Specify number of dimensions
+    '''
+
+    for i, component in enumerate(component_names):
+        print('---------- Normalizing component {} ----------'.format(component))
+        try:
[…]

### 2021-08-04 src/visualization.py

@@ -1,0 +1,232 @@
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+import itertools
+
+
+def linegraph(f_name, type, dims, component_names, colors_list):
+    ''' 
+    Plot the explained variance across dimensions using the model results
+
+    @params:
+        f_name: Name of the file
+        type: Type of reduction method, pca or autoencoders
+        dims: Number of dimensions
+        component_names: Gas/particulate list
+        colors_list: List of colors to be used
+    '''
+    
+    num_of_comp = list(range(2,dims+1))
+    # Plot results
+    plt.figure(figsize=(12,10))
+    for i, component in enumerate(component_names):
+        # AE or PCA
+        if type == 'ae':
+            file_name = '{FILE}{COMPONENT}_ae_results_gs.csv'.format(FILE=f_name, COMPONENT=component)
+            plt_title = 'Autoencoder Reduced Representation of Air Pollutants'
+        elif type == 'pca':
+            file_name = '{FILE}{COMPONENT}_pca_results_gs.csv'.format(FILE=f_name, COMPONENT=component)
+            plt_title = 'PCA Reduced Representation of Air Pollutants'
+        else:
+            print('Type must be "ae" or "pca"')
+            quit()
+
+        # Read in model results
+        model = pd.read_csv(filepath_or_buffer=file_name)
+        variance = model['{}_var'.format(component)]
+        r2 = model['{}_r2'.format(component)]
+        
+
+        plt.plot(num_of_comp, variance[:dims-1], label = '{}'.format(component), linestyle = '-', marker = '+', color = colors_list[i])
+        plt.plot(num_of_comp, r2[:dims-1], linestyle = '-.', marker = 'H', color = colors_list[i])
+        plt.rcParams.update({'font.size': 22})
+        plt.tick_params(axis='both', which='major', labelsize=28)
+       # plt.xlabel('Dimension')
+       # plt.ylabel('% Explained Variance')
+        plt.xlabel('')
+        plt.ylabel('')
+        #plt.title(plt_title)
+        plt.ylim([0,1])
+        plt.legend()
+    plt.show()
+
+
+def scatter(f_name, type, component, color):
+    '''
[…]

### 2021-08-04 visualization.py

@@ -1,201 +1,0 @@
-import pandas as pd
-import numpy as np
-import matplotlib.pyplot as plt
-import seaborn as sns
-import itertools
-
-
-def linegraph(f_name, type, dims, component_names, colors_list):
-    ''' 
-    Plot the explained variance across dimensions using the model results
-
-    @params:
-        f_name: Name of the file
-        type: Type of reduction method, pca or autoencoders
-        dims: Number of dimensions
-        component_names: Gas/particulate list
-        colors_list: List of colors to be used
-    '''
-    
-    num_of_comp = list(range(2,dims+1))
-    # Plot results
-    plt.figure(figsize=(12,10))
-    for i, component in enumerate(component_names):
-        # AE or PCA
-        if type == 'ae':
-            file_name = '{FILE}{COMPONENT}_ae_results_gs.csv'.format(FILE=f_name, COMPONENT=component)
-            plt_title = 'Autoencoder Reduced Representation of Air Pollutants'
-        elif type == 'pca':
-            file_name = '{FILE}{COMPONENT}_pca_results_gs.csv'.format(FILE=f_name, COMPONENT=component)
-            plt_title = 'PCA Reduced Representation of Air Pollutants'
-        else:
-            print('Type must be "ae" or "pca"')
-            quit()
-
-        # Read in model results
-        model = pd.read_csv(filepath_or_buffer=file_name)
-        variance = model['{}_var'.format(component)]
-        r2 = model['{}_r2'.format(component)]
-        
-
-        plt.plot(num_of_comp, variance[:dims-1], label = '{}'.format(component), linestyle = '-', marker = '+', color = colors_list[i])
-        plt.plot(num_of_comp, r2[:dims-1], linestyle = '-.', marker = 'H', color = colors_list[i])
-        plt.rcParams.update({'font.size': 22})
-        plt.tick_params(axis='both', which='major', labelsize=28)
-       # plt.xlabel('Dimension')
-       # plt.ylabel('% Explained Variance')
-        plt.xlabel('')
-        plt.ylabel('')
-        #plt.title(plt_title)
-        plt.ylim([0,1])
-        plt.legend()
-    plt.show()
-
-
-def scatter(f_name, type, component, color):
-    '''
-    Autoencoder scatter plot of multiple components for the first two dimensions
-
[…]

### 2021-08-04 paper-draft/paper.tex
Removed unncecessary files from paper-draft/

@@
 \usepackage[pass, letterpaper]{geometry}
-%\usepackage[margin=1in]{geometry}
+\usepackage{amsmath}
+\usepackage{amsfonts}
 \usepackage{xcolor}
@@
 \section{Results}
-First, we wanted to explore the correlation between days in our dataset. We calculated
-the correlation coefficients and created a correlation matrix of the days that had
-correlation coefficients of the highest magnitude.\par
-
-% pipeline
-\begin{figure}[h]
-\includegraphics[width=\textwidth]{/home/nicks/Documents/paper-draft/images/corr_matrix_25.png}
-\caption{Correlation Matrix}
-\label{fig:pipeline}
-\end{figure}
-
 We compared the percent explained variance across one hundred twenty dimensions for
@@
 all of them were located near the center of the clustering. Conversely, we found that many 
-of the outliers were smaller cities in rural locations.
+of the outliers were smaller cities in rural locations. \par
 
+\newpage
+
+To further explore the meaning of the hidden dimensions, we calculated the Pearson correlation
+coefficient of the normalized timeseries data and the embeddings. The Pearson correlation
+coefficient allows us allows us to observe the strength of the linear correlation between
+the encoded and unencoded values. 
+\begin{equation*}
+    r_{XY}=\frac{cov(X,Y)}{\sigma_{X} \cdot \sigma_{Y}}
+\end{equation*}
+\textit{Where cov(X,Y) is the covariance between X and Y \\
+and $\sigma_{X},\sigma{Y}$ are the standard deviations of X and Y respectively}
+
+\vspace{0.1in}
+
+We calculated the correlation coefficients and created a correlation matrix of the days that had
+correlation coefficients of the highest magnitude.\par
+
+% pipeline
+\begin{figure}[h]
+\includegraphics[width=\textwidth]{/home/nicks/Documents/paper-draft/images/corr_matrix_25.png}
+\caption{Correlation Matrix}
+\label{fig:matrix}
+\end{figure}
 \section{Discussion}

### 2021-09-08 src/autoencoder.py
Updated code

@@ -129,7 +129,7 @@ def ae_run(dims, component):
     
     @params:
         dims: Number of dimensions
-        component_names: Gas/particulate list
+        component: Gas/particulate 
     '''
     
     # Open normalized data
@@ -142,8 +142,8 @@ def ae_run(dims, component):
     param_grid = pd.read_csv(f'/home/nicks/github_repos/Pollution-Autoencoders/data/grid_params/hyperparams.csv')
     
     # Write headers for model metrics and vecs; erase previous model
-    metrics_file = f'/home/nicks/github_repos/Pollution-Autoencoders/data/model_metrics/{component}_metrics.csv'
-    vec_file = f'/home/nicks/github_repos/Pollution-Autoencoders/data/vec/{component}_vec.csv'
+    metrics_file = f'/home/nicks/github_repos/Pollution-Autoencoders/data/model_metrics/{component}_metrics_ten.csv'
+    vec_file = f'/home/nicks/github_repos/Pollution-Autoencoders/data/vec/{component}_vec_ten.csv'
     
     # Set x as the normalized values, y as the daily average of final day
     x = dfx.values
@@ -157,21 +157,24 @@ def ae_run(dims, component):
     lr = param_grid['lr'][0]
     batch = param_grid['batch'][0]
     epochs = param_grid['epochs'][0]
+    # List of key dimensions to advance counter
+    param_dims = list(param_grid['dim'])
     counter=0
 
     vector = []
     var_list = []
     r2_list = []
-
+    
     for dim in num_of_dims:
         # Change to next set of hyperparams every 10 dims
         print(f'lr: {lr} batch: {batch} epochs: {epochs}')
-        if dim%10==0 and dim<120:
-            counter+=1
+
+        if dim in param_dims:
             lr = param_grid['lr'][counter]
             batch = param_grid['batch'][counter]
             epochs = param_grid['epochs'][counter]
-            
+            counter+=1
+        
         # Define the linear regression model
         regr = LinearRegression()
         print('---------- Autoencoder dim {} for {} ----------'.format(dim, component))
@@ -179,8 +182,8 @@ def ae_run(dims, component):
         input_data = Input(shape=(191,))
 
         # Create dense AE layers
[…]

### 2021-09-08 src/visualization.py

@@ -17,16 +17,18 @@ def linegraph(f_name, type, dims, component_names, colors_list):
         colors_list: List of colors to be used
     '''
     
-    num_of_comp = list(range(2,dims+1))
+    num_of_comp = list(range(2,dims+1)) # TODO: delete this later
+    num_of_dims = list(range(1, dims+1))
     # Plot results
     plt.figure(figsize=(12,10))
     for i, component in enumerate(component_names):
         # AE or PCA
         if type == 'ae':
-            file_name = '{FILE}{COMPONENT}_ae_results_gs.csv'.format(FILE=f_name, COMPONENT=component)
+            #file_name = f'{f_name}{component}_metrics.csv'
+            file_name = '/home/nicks/github_repos/Pollution-Autoencoders/data/model_metrics/test_metrics.csv'
             plt_title = 'Autoencoder Reduced Representation of Air Pollutants'
         elif type == 'pca':
-            file_name = '{FILE}{COMPONENT}_pca_results_gs.csv'.format(FILE=f_name, COMPONENT=component)
+            file_name = f'{f_name}{component}_metrics.csv'
             plt_title = 'PCA Reduced Representation of Air Pollutants'
         else:
             print('Type must be "ae" or "pca"')
@@ -34,8 +36,8 @@ def linegraph(f_name, type, dims, component_names, colors_list):
 
         # Read in model results
         model = pd.read_csv(filepath_or_buffer=file_name)
-        variance = model['{}_var'.format(component)]
-        r2 = model['{}_r2'.format(component)]
+        variance = model['variance']
+        r2 = model['r2']
         
 
         plt.plot(num_of_comp, variance[:dims-1], label = '{}'.format(component), linestyle = '-', marker = '+', color = colors_list[i])
@@ -217,16 +219,17 @@ def correlation(X_data_file, Y_data_file, component):
     plt.show()
     '''
 
-### RUN ###
+### Function Calls ###
 
 COMPONENT_NAMES = ['co', 'no', 'no2', 'o3', 'so2', 'pm2_5', 'pm10', 'nh3']
 #COMPONENT_NAMES = ['co']
[…]

### 2021-09-08 paper-draft/paper.tex
Updated code

@@
 \title{Using a Deep Autoencoder Model to Characterize Air Pollution}
-\author{Nicolas Shannon$^{1}$, Bishnu Timalsena$^{2}$, Raul Chavez Jr.$^{3}$, \\Vandana Nunna Lakshmi$^{2}$}
+\author{Nicolas Shannon$^{1}$, Vandana Nunna Lakshmi$^{2}$}
 \date{}
@@
 {$^{2}$Department of Computer Science, University North Texas, Denton, TX, US}\\
-{$^{3}$Department of Electrical and Computer Engineering, University of Texas 
-    - Rio Grande Valley, Brownsville, TX, US \note{$\leftarrow$ fix formatting}}
+%{$^{3}$Department of Electrical and Computer Engineering, University of Texas 
+%    - Rio Grande Valley, Brownsville, TX, US \note{$\leftarrow$ fix formatting}}
 \end{centering}
@@
 \subsection{A Deep Learning Approach}
-In this paper we compare the autoencoder model with principle component analysis (PCA), a much
-older technique for dimensional reduction. Using a deep autoencoder has several advantages.
-First, deep autoencoders are better at compressing highly complex data into few dimensions.
-Autoencoders with a nonlinear encoder and decoder function can learn more powerful nonlinear
-generalizations than PCA. \parencite{Goodfellow16}.
-\par \note{Autoencoder grapic from poster here}
+In this paper we employ autoencoder and  principle component analysis (PCA) models to 
+feed the pollutant data and perform data compression. The resulting embeddings are 
+compared in terms of effeciency and data loss. The features are encoded at different
+scales of dimensions to check the fluctuation in explainable variance when predicting a 
+given target. Given that each feature in the dataset represents the daily average of a 
+specific pollutant in that location and that pollution levels do not vary drastrically 
+on a day-to-day basis, it is evident that there is a strong correlation among features
+\par While a healthy correlation among features is encouraged, abnormally strong values
+relfect redundancy and end up being a burden on the model.
+\parencite{featureredundancy}. First, deep autoencoders are better at compressing highly 
+complex data into fewer dimensions. Autoencoders with a nonlinear encoder and decoder 
+function can learn more powerful nonlinear generalizations than PCA. 
+\parencite{Goodfellow16}.
 
 \section{Related Work}
-The idea of generating embeddings for a location is a highly potential expanse for research.
-The two major forces behind its competency are, the fact that large volumes of data is being
-effectively compressed and that these embeddings, when plugged into any model could effectively
-represent any given location.  A lot of research is being done on location embeddings in recent
-times. For instance, location embeddings are being generated based on human-mobility data 
-(Crivellari, A.; 2019) to understand the behavioral proximity between different locations 
-irrespective of their spatial distances. The sequence of locations that are frequently
-travelled were fed to a Skip-gram Word2vec model to generate the location embeddings. Another 
-approach that was implemented to leverage location embeddings is, GPS2Vec ( Yifang Yin, 2019).
-This application consists of a neural network that would generate embeddings for a location, 
-based on the semantic contexts that are sourced from the user content published on multiple 
-digital platforms. With the integration of these embeddings, a successful geo-tagged image
-classification task is demonstrated. In another effort (Carl Yang, 2019), the location 
-embeddings derived from user content such as addresses, phone numbers, images, place names and 
-all other details from multiple individual online posts, are compressed to form low dimensional
-embeddings to solve the place duplication issue. Deduplication of locations, is a paramount 
-task for any digital platform since they provide a place graph to its users, which is the 
-result of merging location data from multiple sources. This merge leads to duplicity since 
-different sources could hold different names for the same place. By applying K-NN search, 
-pair-wise duplication prediction on these generated location embeddings, the application can
-identify the duplicates and eliminate them to result in an efficient place graph.
+The idea of generating embeddings for a location is a highly potential expanse for
+research. The two major forces behind its competency are, the fact that large volumes of
+data is being effectively compressed and that these embeddings, when plugged into any
+model could effectively represent any given location.  A lot of research is being done 
+on location embeddings in recent times. For instance, location embeddings are being 
+generated based on human-mobility data \parencite{GeoEmbeddings} to understand the 
+behavioral proximity between different locations irrespective of their spatial distances.
+The sequence of locations that are frequently travelled were fed to a Skip-gram Word2vec 
+model to generate the location embeddings. Another approach that was implemented to 
+leverage location embeddings is, GPS2Vec  \parencite{GPS2Vec}.  This application consists
+of a neural network that would generate embeddings for a location, based on the semantic 
+contexts that are sourced from the user content published on multiple digital platforms.
+With the integration of these embeddings, a successful geo-tagged image classification 
+task is demonstrated. In another effort  \parencite{PlaceDeduplication}, the location 
+embeddings derived from user content such as addresses, phone numbers, images, place
+names and all other details from multiple individual online posts, are compressed to 
+form low dimensional embeddings to solve the place duplication issue. Deduplication of 
+locations, is a paramount task for any digital platform since they provide a place graph
+to its users, which is the result of merging location data from multiple sources. This 
+merge leads to duplicity since different sources could hold different names for the same 
+place. By applying K-NN search, pair-wise duplication prediction on these generated 
+location embeddings, the application can identify the duplicates and eliminate them to 
+result in an efficient place graph.\par
+In 2018, a new approach, “DeepMove” is proposed \parencite{DeepMove}. This 
+implementation employs Trip model and OD model to learn latent representations of 
+places.  The approach is successfully applied to New York city yellow taxi dataset – 
+“includes pick-up/drop-off times, the number of passengers, pick-up/drop-off locations 
+represented by longitude and latitude, trip distance, rate types, payment type, and
+itemized fares” \parencite{DeepMove}. Similarities among locations in New York city are 
+derived effectively with “DeepMove”. To resolve the dissonance that arises from the 
+fusion of geographic and semantic features of a location, an unsupervised learning 
+implementation had been demonstrated that would derive location embeddings from human 
+trajectories \parencite{DisentangledLocationEmbeddings}. With the ever- growing 
+footprint of digital and social media platforms in everyday life of an individual, a
+lot of research is happening to generate embeddings based on the content posted in 
+different locations. LeGo-CM is one such effort that would generate embeddings of 
+geotagged tweets \parencite{GeotaggedTweets}. These resultant embeddings could be 
+applied to understand the kind of content each location reflects.Venue2Vec is another 
+idea that is implemented to enhance the location-aware services in various platforms 
+\parencite{Venue2Vec}. This model is trained on semantic data, temporal-spatial context 
+and sequential relations among locations to generate embeddings. This makes sure that 
+locations with similar data would fall close, when plotted in a multi-dimensional space,
+based on the input features. Hence, the model is effective in predicting the next and  
+future check-in location. Community Flow prediction is another area of interest where
+embeddings are a major problem solver. Geo-contextual Multitask Embedding Learner (GMEL)
+, is a model that uses “Graph Attention Network” to encode geographical contexts and 
+spatial correlations to generate location embeddings \parencite{CommutingFlowPrediction}.
+These embeddings are further used to predict the community fow that serves as a vital
+factor in policy establishment and metro planning.
 
@@
 We used two different methods to reduce the dimensionality of our data - Principle Component
-Analysis and Undercomplete Autoencoders. A simple linear regression was performed on 
+Analysis and undercomplete autoencoders. A simple linear regression was performed on 
 the encoded data that was generated by both the PCA and AE models. Additionally, we performed
@@
 and the air quality index, the latter of which we did not end up using.\par
-We initially looked at data for the eight pollutants over a six-and-a-half month time 
-frame: January 27th, 2020 through July 6th, 2021. However, we found that we achieved
-better results with just four months of of data for our input features. Our final 
+We looked at data for the eight pollutants over a six-and-a-half month time 
+frame: January 27th, 2020 through June 6th, 2021. Our final 
 dataset included just over eighteen thousand cities in the United States and other U.S. 
-territories. Additionally, we found it easier to focus on a single gas when developing 
+territories. We found it easier to focus on a single gas when developing 
 the models, so we chose to concentrate on CO during the engineering process. Given
 enough time and computational power, the techniques performed on the CO model would be 
-easily extensible and transferable to the other seven pollutants.
+easily extensible and transferable for the other seven pollutants.
 
@@
         \hline
-        Time period  & 6.5 months    \\
-        Time frequency & Daily  \\
+        Sampling period  & 11/27/20 to 06/05/21\\
+        Time interval & Daily  \\
         $\#$ of cities & 18,526 \\
-        $\#$ of features & 120 (originally 190) \\
+        $\#$ of features & 191 \\
         $\#$ of component gases & 8 \\
-        Dependent variable & Single day of time series (3/27/21) \\
+        Dependent variable & Single day of time series (06/05/21) \\
         \hline
@@
 OpenWeather's Pollution API ordinarily returns hourly values, but we opted to use the daily
-average pollution levels as our feature set. We resolved to feature engineer the daily 
-averages and to reduce the time scale to four months in order to keep the data 
-proportional to the number of features and thus avoid the curse of dimensionality. 
+average pollution levels as our feature set in order to make the number of starting dimensons
+more manageable. We resolved to feature engineer the daily 
+averages and proportional to the number of features and thus avoid the curse of dimensionality. 
 \parencite{Trunk79}. \note{Idk about this} \par
@@
 
-\subsection{Autoencoder and PCA Models}
+\subsection{Undercomplete Autoencoder Model}
 The autoencoder model used consists of an encoded and decoded layer. We performed
@@
 \end{itemize}
-We used the parameters that appeared most frequently over the one hundred twenty dimensions and 
-applied them over the entire range so as not to overfit.\note{$\leftarrow$ not sure if this is correct}
+K-folds cross validation was used to validate the grid search using five folds. We selected the
+highest coefficient of determination that appeared in each of the five folds for every given
+dimension. The most frequently occuring combination of hyperparameters was then used to create
+the finalized embedding.
 
@@
         \hline
-        Activation function & Leaky ReLU \\
+        Activation function & tanh\\
         Loss function & Mean Squared Error  \\
@@
 
+\subsection{Principle Component Analysis}
 The PCA model was trained with k-fold cross-validation using five folds. The dataset was
 split into a training, validation, and hold-out test set.
+
+\subsection{Model Validation}
+K-folds cross validation was performed on the linear regression test using five folds. The folds
+with the best and worst variance scores were extracted so as to make an equitable evaluation of
+the model.
 

### 2021-09-18 paper-draft/paper.tex
Overhauled ae to include validation split, optimized run function

@@
 \begin{figure}[h]
-\includegraphics[width=\textwidth]{/home/nicks/Documents/paper-draft/images/co-concentration-table.png}
+\includegraphics[width=\textwidth]{/home/nicks/github_repos/Pollution-Autoencoders/paper-draft/images/co-concentration-table.png}
     \caption{Carbon Monoxide Exposure Table, Parts Per Million (ppm)}
@@
 \begin{figure}[h]
-\includegraphics[width=\textwidth, height=6cm]{/home/nicks/Documents/paper-draft/images/pipeline.png}
+\includegraphics[width=\textwidth, height=6cm]{/home/nicks/github_repos/Pollution-Autoencoders/paper-draft/images/pipeline.png}
 \caption{Data Pipeline}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nicks/Documents/paper-draft/images/ae_all_dims.png} 
+    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nicks/github_repos/Pollution-Autoencoders/paper-draft/images/ae_all_dims.png} 
     \caption{Autoencoder Dimensionality Reduction}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nicks/Documents/paper-draft/images/ae_all_dims.png}
+    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nicks/github_repos/Pollution-Autoencoders/spaper-draft/images/ae_all_dims.png}
     \caption{PCA Dimensionality Reduction}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nicks/Documents/paper-draft/images/co_scatter_ae_wlist.png} 
+    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nicks/github_repos/Pollution-Autoencoders/paper-draft/images/co_scatter_ae_wlist.png} 
     \caption{Outlier Cities}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nicks/Documents/paper-draft/images/co_scatter_ae_large_cities.png}
+    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nicks/github_repos/Pollution-Autoencoders/paper-draft/images/co_scatter_ae_large_cities.png}
     \caption{Densely Populated Cities}
@@
 \begin{figure}[h]
-\includegraphics[width=\textwidth]{/home/nicks/Documents/paper-draft/images/corr_matrix_25.png}
+\includegraphics[width=\textwidth]{/home/nicks/github_repos/Pollution-Autoencoders/paper-draft/images/corr_matrix_25.png}
 \caption{Correlation Matrix}

### 2021-09-18 src/autoencoder.py

@@ -3,6 +3,7 @@ import pandas as pd
 import itertools
 import csv
 import os
+import linreg # src/linreg.py
 from tensorflow.keras.models import Model
 from tensorflow.keras.layers import Dense, Input
 from tensorflow.keras.optimizers import Adam
@@ -12,148 +13,73 @@ from sklearn.model_selection import KFold, train_test_split
 from sklearn.metrics import r2_score
 
 
-def autoencoder(dfx, y, dim, component, activation, lr=0.0001, batch=128, epochs=50):
+def autoencoder(dim, X_train, Y_train, component, activation, lr, batch, epochs):
     '''
     Autoencoder model
     
     @params:
-        dfx: DataFrame of normalized x values
-        y: Dependent variable to be predicted
         dim: Current dimension
+        X_train, Y_train: training data
         component: Name of gas or particulate
         activation: Activation function represented as a tuple pair
-        lr: Learning rate, default of 0.0001
-        batch: Batch size, default of 128
-        epochs: Number of epochs, default of 50
+        lr: Learning rate
+        batch: Batch size
+        epochs: Number of epochs
     @return:
-        variance_list: List of explainable variance scores
-        r2_list: List of R squared scores
-        encoded_data: The encoded X training data
+        encoded_layer: The encoded layer responsible for encoding the data
+        encoded_train_data: Training embeddings needed to fit the regression model
+        Y_train: The reduced Y training data to match size of the encoded training data
     '''
 
-    # Define the linear regression model
-    regr = LinearRegression()
-
-    print('---------- Autoencoder dim {} for {} ----------'.format(dim, component))
-    # Training and test splits
-    train, dev, train_labels, dev_labels = train_test_split(dfx, y, test_size=0.20, random_state=40)
-    train, test, train_labels, test_labels = train_test_split(train, train_labels, test_size=0.20, random_state=40)
+    # Additional split of train/test data; keep dev set for validation
+    X_train, X_dev, Y_train, Y_dev = train_test_split(X_train, Y_train, test_size=0.20, random_state=40)
     input_data = Input(shape=(191,))
 
-    x_train = train.loc[:, train.columns]
-    x_test = test.loc[:, test.columns]
-    x_dev = dev.loc[:, dev.columns]
-
+    # Create layers
[…]

### 2021-09-18 src/linreg.py

@@ -1,175 +1,36 @@
 import numpy as np
 import pandas as pd
-import matplotlib.pyplot as plt
-import os
-import sys
-import operator
-from numpy.core.numeric import NaN
-from tensorflow.keras.models import Model
-from tensorflow.keras.layers import Dense, Input
-from tensorflow.keras.optimizers import Adam
-from sklearn.preprocessing import Normalizer
 from sklearn.linear_model import LinearRegression
-from sklearn.model_selection import KFold, train_test_split
 from sklearn.metrics import r2_score
 
 
-def linreg(X_train, X_test, Y_train, Y_test, component, folds=0, dev=False):
-    '''
-    Simple linear regression
-
+def regression(train_data, test_data, train_labels, test_labels):
+    ''' 
+    Perform a linear regression 
+    
     @params:
-        dfx: DataFrame of normalized x values
-        y: Predicted value
-        component: Name of gas or particulate
-    '''
-
+        train_data: Encoded data used to fit the regression
+        test_data: Encoded data used for prediction
+        train_labels, test_labels: testing/training labels
+    @return:
+        variance, r2: The variance/r2 scores of the encoded test set
+    '''    
+    
+    # Define the regression model
     regr = LinearRegression()
+
+    #print('size of trained data', len(train_data))
+    #print('size of trained labels', len(train_labels))
+    # Fit the regression model with the encoded training data
+    regr.fit(train_data, train_labels)
     
-    input_data = Input(shape=(191,)) # <- Not sure if needed
+    Y_pred = regr.predict(test_data) 
     
-    # Linear regression on validation/train set
-    if dev == True:
-        print(f'Linear regression validation -- fold {folds} -- component {component}')
-        train_reg, dev_reg, train_labels_reg, dev_labels_reg = train_test_split(X_train, Y_train, test_size=0.33, random_state=42)
-        regr.fit(train_reg, train_labels_reg)
-        res_sum_square= np.mean((regr.predict(dev_reg) - dev_labels_reg ** 2))
-        var_val=regr.score(dev_reg, dev_labels_reg)
-        Y_test_predict = regr.predict(dev_reg)
-        r2_val=r2_score(dev_labels_reg, Y_test_predict)
-        return (regr.intercept_, regr.coef_, res_sum_square, var_val, r2_val)
+    # Variance and r2 scores for the regression model
[…]

### 2021-09-18 src/pca.py

@@ -11,7 +11,7 @@ from sklearn.preprocessing import Normalizer
 from sklearn.cluster import KMeans
 
 
-def pca(i, X_train, X_test, Y_train, Y_test, folds, component, dev=False):
+def pca(i, X_train, X_test, Y_train, Y_test, component, dev=False):
     '''
     Perform Linear Regression on either train/validation set or on test set for PCA
 
@@ -27,12 +27,12 @@ def pca(i, X_train, X_test, Y_train, Y_test, folds, component, dev=False):
 
     # Encode data
     encoded_X_train = pca.fit_transform(X_train)
-    encoded_X_test = pca.transform(X_test)
+    #encoded_X_test = pca.transform(X_test)
     
     # Validate the data using an embedding
     print(f'\nLinear regression -- component {component}')
     regr.fit(encoded_X_train, Y_train)
-    #res_sum_square = np.mean((regr.predict(encoded_X_test) - Y_test ** 2))
+
     Y_test_predict = regr.predict(encoded_X_test)
 
     variance = regr.score(encoded_X_test, Y_test)
@@ -40,32 +40,16 @@ def pca(i, X_train, X_test, Y_train, Y_test, folds, component, dev=False):
 
     return (variance, r2, encoded_X_test)
     
-    # TODO: Remove code below (maybe?)
-    '''
-    # Linear regression on validation/train set
-    if dev == True:
-        if i%10 == 0:
-            print('PCA validation -- fold {} -- component {}'.format(folds,i))
-        train_pca, dev_pca, train_labels_pca, dev_labels_pca = train_test_split(X_train, Y_train, test_size=0.33, random_state=42)
-        model.fit(train_pca, train_labels_pca)
-        res_sum_square= np.mean((model.predict(dev_pca) - dev_labels_pca ** 2))
-        var_val=model.score(dev_pca, dev_labels_pca)
-        Y_test_predict = model.predict(dev_pca)
-        r2_val=r2_score(dev_labels_pca, Y_test_predict)
-        return (model.intercept_, model.coef_, res_sum_square, var_val, r2_val)
-
-    # Linear regression on test set
-    else:
-        if i%10 == 0:
-            print('PCA Test -- component {}'.format(i))
-        model.fit(X_train, Y_train)
-        res_sum_square= np.mean((model.predict(X_test) - Y_test ** 2))
-        var_val=model.score(X_test, Y_test)
-        Y_test_predict = model.predict(X_test)
-        r2_val=r2_score(Y_test, Y_test_predict)
[…]

### 2021-09-18 src/visualization.py

@@ -54,6 +54,59 @@ def linegraph(f_name, type, dims, component_names, colors_list):
     plt.show()
 
 
+def metrics_comparison(type, dims, component_names, colors_list):
+    ''' 
+    Plot the explained variance across dimensions using the model results
+
+    @params:
+        type: Type of reduction method, pca or autoencoders
+        dims: Number of dimensions
+        component_names: Gas/particulate list
+        colors_list: List of colors to be used
+    '''
+     
+    num_of_comp = list(range(2,dims+1)) # TODO: delete this later
+    num_of_dims = list(range(1, dims+1))
+    # Plot results
+    plt.figure(figsize=(12,10))
+    for i, component in enumerate(component_names):
+        # AE or PCA
+        if type == 'ae':
+            high_file = f'/home/nick/github_repos/Pollution-Autoencoders/data/test_metrics/ae/derived/{component}_best_metrics.csv'
+            low_file = f'/home/nick/github_repos/Pollution-Autoencoders/data/test_metrics/ae/derived/{component}_worst_metrics.csv'
+            plt_title = 'Autoencoder Reduced Representation of Air Pollutants'
+        elif type == 'pca':
+            high_file = f'/home/nick/github_repos/Pollution-Autoencoders/data/test_metrics/pca/derived/{component}_best_metrics.csv'
+            low_file = f'/home/nick/github_repos/Pollution-Autoencoders/data/test_metrics/pca/derived/{component}_worst_metrics.csv'
+            plt_title = 'PCA Reduced Representation of Air Pollutants'
+        else:
+            print('Type must be "ae" or "pca"')
+            quit()
+
+        # Read in model results
+        best = pd.read_csv(filepath_or_buffer=high_file)
+        worst = pd.read_csv(filepath_or_buffer=low_file)
+        # Best and worst scores read in
+        best_variance = best['variance']
+        best_r2 = best['r2']
+        worst_variance = worst['variance']
+        worst_r2 = worst['r2']
+
+        # Plots
+        plt.plot(num_of_comp, best_variance[:dims-1], label = f'{component} high variance', linestyle = '-', marker = '+', color = 'green')
[…]

### 2021-09-18 src/autoencoder.py
Overhauled pca to follow same structure as ae

@@ -69,11 +69,13 @@ def ae_train_test(dims, X_train, X_test, Y_train, Y_test, component, param_grid)
     
     @params:
         dims: Number of dimensions to train
+        X_train, X_test, Y_train, Y_test: train/test data used for model
         component: Gas/particulate
         param_grid: List of key hyperparameters to test 
-        production: Boolean to determine whether to train/test or to create full embedding
     '''
 
+    # Metrics file
+    file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/ae/{component}_metrics'
     # Train Autoencoder model
     variance_list = []
     r2_list = []
@@ -119,7 +121,7 @@ def ae_train_test(dims, X_train, X_test, Y_train, Y_test, component, param_grid)
     # Write variance and r2 scores of each gas for every dimension 
     output_dict = {'dim': num_of_dims, 'variance' : variance_list, 'r2' : r2_list}
     metrics_data = pd.DataFrame(data=output_dict)
-    metrics_data.to_csv(path_or_buf=metrics_file, index=False)
+    metrics_data.to_csv(path_or_buf=file_name, index=False)
 
 
 def ae_run(dim, X, X_train, Y_train, lr, batch, epochs, component, cities):
@@ -129,6 +131,8 @@ def ae_run(dim, X, X_train, Y_train, lr, batch, epochs, component, cities):
      
     @params:
         dim: Dimension to embed
+        X: Normalized data to encode
+        X_train, Y_train: Train data used in model creation
         component: Gas/particulate 
         cities: List of cities to append to the embedding
         param_grid: The correct hyperparameters to be used. Note the dimensions that were tested in the 
@@ -271,7 +275,7 @@ def main():
     X_train, X_test, Y_train, Y_test = train_test_split(X, Y, test_size=0.20, random_state=40)
 
     ### Function calls ###
-    
+
     #grid_search(X, Y, folds, 'co', iter_dims, key_params)
     #ae_train_test(dims, X_train, X_test, Y_train, Y_test, component_test, param_grid)
     #ae_run(dims, X, X_train, Y_train, 0.01, 64, 75, component_test, cities)

### 2021-09-18 src/pca.py

@@ -1,8 +1,8 @@
 import pandas as pd
 import numpy as np
-import matplotlib.pyplot as plt
 import csv
-
+import os
+import linreg # src/linreg.py
 from sklearn.decomposition import PCA
 from sklearn.model_selection import KFold, train_test_split
 from sklearn.metrics import r2_score
@@ -10,62 +10,37 @@ from sklearn.linear_model import LinearRegression
 from sklearn.preprocessing import Normalizer
 from sklearn.cluster import KMeans
 
-
-def pca(i, X_train, X_test, Y_train, Y_test, component, dev=False):
+def pca(i, X_train, Y_train, component):
     '''
     Perform Linear Regression on either train/validation set or on test set for PCA
 
     @params:
-        i: number of components
-        X_train, X_test, Y_train, Y_test: testing/training lists
-        dev: set to true if doing validation testing; false if using test set data      
+        i: Number of components
+        X_train, Y_train: Training data
+        component: Gas/particulate 
     '''
 
-    # Define the models
+    # Define the model and fit the trained embedding
     pca = PCA(n_components=i)
-    regr = LinearRegression()
-
-    # Encode data
     encoded_X_train = pca.fit_transform(X_train)
-    #encoded_X_test = pca.transform(X_test)
     
-    # Validate the data using an embedding
-    print(f'\nLinear regression -- component {component}')
-    regr.fit(encoded_X_train, Y_train)
+    return pca, encoded_X_train
 
-    Y_test_predict = regr.predict(encoded_X_test)
 
-    variance = regr.score(encoded_X_test, Y_test)
-    r2 = r2_score(Y_test, Y_test_predict)
-
-    return (variance, r2, encoded_X_test)
-    
-def linreg(encoded_train_data, X_train, X_test, Y_train, Y_test):
-    ''' 
-    Perform a linear regression 
-    
-    @params:
-        encoded_train_data: Data used to fit the regression
-        X_train, X_test, Y_train, Y_test: testing/training lists
-    @return:
-        variance, r2: The variance/r2 scores of the encoded test set
-    '''  
-
-def pca_train(x, y, folds, dims, component):
+def pca_train_test(dims, x, y, folds, component):
     ''' 
     Run PCA using k-fold cross validation strategy
 
     @params:
-        x: Normalized data to encode
-        y: Dependent variable
+        dims: Number of dimensions to train
+        x, y: Independent/dependent data used for k-fold split
[…]

### 2021-12-21 paper-draft/paper.tex
Maintenance commit

--- 2021-09-08 paper-draft/paper.tex
@@
 \newcommand\note[1]{\textbf{\textcolor{red}{#1}}}
+\newcommand{\comment}[1]{}
 
@@
 \captionsetup{justification = raggedright, singlelinecheck = false, labelfont=bf}
+% To cite, use \parencite{} or \textcite{}
 
-\title{Using a Deep Autoencoder Model to Characterize Air Pollution}
+\title{Using an Autoencoder Model to Characterize Air Pollution}
 \author{Nicolas Shannon$^{1}$, Vandana Nunna Lakshmi$^{2}$}
 \date{}
-% To cite, use \parencite{} or \textcite{}
 
 \begin{document}
+
+% Indentation
+\setlength{\parindent}{3ex}
+\setlength{\parskip}{0.25ex}
 
@@
 {$^{2}$Department of Computer Science, University North Texas, Denton, TX, US}\\
-%{$^{3}$Department of Electrical and Computer Engineering, University of Texas 
-%    - Rio Grande Valley, Brownsville, TX, US \note{$\leftarrow$ fix formatting}}
 \end{centering}
@@
 \begin{abstract}
-Air pollutants are both a cause and a symptom of climate change. Global warming has the potential to raise
-the levels of various air pollutants. The rise in surface air pollutants can lead to heat trapped close to 
-the Earth's surface by virtue of the greenhouse effect, which further perpetuates the cycle. In order to
-more effectively combat climate change, it is important to understand and characterize the source of 
-emissions. We developed a technique to characterize cities by the levels of various air toxins using an
-undercomplete autoencoder model. By representing the pollution levels of eighteen thousand cities across
-the United States, we explore some of the more prominent features that characterize urban pollution.
-We compare two techniques for reducing the input dimensionality - Principle Component Analysis (PCA) and
-Undercomplete Autoencoders.
+Climate change is one of the most complex challenges humanity has ever faced. It is a
+multi-dimensional problem that poses economic, social, scientific, political, and 
+moral questions that must be addressed. One of the most important ways of 
+combating climate change is to research pollution mitigation strategies 
+so that policy makers and government planners are better equipped to  make 
+informed decisions. We focused on the principle cause of global warming:
+the polluting gases responsible for the rapid shift in global climate.
+We characterize cities by the amount of pollution present using an undercomplete autoencoder
+model. By representing the level of eight toxic gases present in numerous cities
+across the United States, we explore some of the more prominent features 
+that characterize urban pollution. We compare two techniques for reducing 
+the dimensionality of our input - Principle Component Analysis (PCA) and
+Autoencoders. The first two hidden dimensions represented
+a large percentage of the explained variance in the model, so we cluster
+the cities in an effort to extract any useful features. 
 \end{abstract}
-
 \vspace{0.25cm}
 \section{Introduction}
-Over the past several decades, air pollution has been a major concern for human health
+
+\par Over the past several decades, air pollution has been a major concern for human health
 and the environment. The EPA sets national air quality standards for  six major air
 toxins: ground level Ozone (O$_3$), Carbon Monoxide (CO), Sulfur Dioxide (SO$_2$),
-Lead (Pb), Nitrogen Dioxide (NO$_2$), and particulate matter (PM). \parencite{NAAQS09}.
-The sources of these emissions vary, but the two largest contributors are industrial 
-processes and motorized vehicles.
+Lead (Pb), Nitrogen Dioxide (NO$_2$), and particulate matter (PM).
+The sources of these emissions vary, but the two largest contributors are industrial processes and motorized vehicles.
+\parencite{NAAQS09}.
+\par We wanted to explore how various polluting gases can be characterized with respect to location in an effort to provide researchers and policy makers with the context and tools necessary to make informed decisions.
+By creating an embedding of time series pollution data for numerous cities throughout the United States, we explore the more interesting features of the model as well as apply it to tangential areas such as the pollution generated by industrial farming and animal husbandry activities.
 
-\newpage
+\par In this paper we use two different methods for data compression, autoencoders and principle component analysis (PCA).
+The models are trained using time series data for eight different polluting gases.
+The trained models are compared in terms of efficiency and data loss.
+The fluctuations in explainable variance when predicting a target allow us to quantify the discrepency between the model results and actual data.
+In this case, the models are trained to predict a single feature out of the time series data, which represents the daily pollutant averages for a single day.
+Once the training process is complete, the trained models are employed to create embeddings for various encoder dimensions.
 
-Short and long term exposure to air toxins pose significant risk to human health. 
-A study by \textcite{Wang19} found a correlation  between increased ambient CO
-levels and outpatient visits for respiratory, cardiovascular, genitourinary, and gastrointenstinal
-diseases. Several studies performed by the EPA also suggest that people with pre-existing health
-conditions are at higher risk when exposed to CO over long periods.\par
-
-\subsection{Air Toxins}
-We wanted to explore how CO and other gases can be characterized with respect to location
-in an effort to provide researchers and policy makers with the context and tools necessary to
-make informed decisions. There are a litany of technologies and tools for pollution prevention
-and control, but most of these options require contextual information and large amounts of data.
-\par
-By creating an embedding of time series pollution data for numerous cities throughout the U.S.,
-we can explore the more interesting features of the model as well as apply it to tangential 
-areas such as the pollution generated by industrial farming and animal husbandry activities.
-
-\begin{figure}[h]
-\includegraphics[width=\textwidth]{/home/nicks/Documents/paper-draft/images/co-concentration-table.png}
-    \caption{Carbon Monoxide Exposure Table, Parts Per Million (ppm)}
-    \label{fig:co-exposure}
-\end{figure}
-
-\subsection{A Deep Learning Approach}
-In this paper we employ autoencoder and  principle component analysis (PCA) models to 
-feed the pollutant data and perform data compression. The resulting embeddings are 
-compared in terms of effeciency and data loss. The features are encoded at different
-scales of dimensions to check the fluctuation in explainable variance when predicting a 
-given target. Given that each feature in the dataset represents the daily average of a 
-specific pollutant in that location and that pollution levels do not vary drastrically 
-on a day-to-day basis, it is evident that there is a strong correlation among features
-\par While a healthy correlation among features is encouraged, abnormally strong values
+\par Given the relatively short time span of our features along with the fact that pollution levels do not tend to vary drastrically day-to-day, it is evident that there is a strong correlation among features.
+While a healthy correlation among features is encouraged, abnormally strong values
 relfect redundancy and end up being a burden on the model.
-\parencite{featureredundancy}. First, deep autoencoders are better at compressing highly 
-complex data into fewer dimensions. Autoencoders with a nonlinear encoder and decoder 
-function can learn more powerful nonlinear generalizations than PCA. 
-\parencite{Goodfellow16}.
+\parencite{featureredundancy}.
+Ideally there would be high correlation between features and the problem class and a low correlation among the features themselves.
 
 \section{Related Work}
-The idea of generating embeddings for a location is a highly potential expanse for
+\par The idea of generating embeddings for a location is a highly potential expanse for
 research. The two major forces behind its competency are, the fact that large volumes of
@@
 model could effectively represent any given location.  A lot of research is being done 
-on location embeddings in recent times. For instance, location embeddings are being 
+on location embeddings in recent times. 
+\par For instance, location embeddings are being 
 generated based on human-mobility data \parencite{GeoEmbeddings} to understand the 
@@
 The sequence of locations that are frequently travelled were fed to a Skip-gram Word2vec 
-model to generate the location embeddings. Another approach that was implemented to 
-leverage location embeddings is, GPS2Vec  \parencite{GPS2Vec}.  This application consists
+model to generate the location embeddings. 
+\par Another approach that was implemented to 
+leverage location embeddings is GPS2Vec  \parencite{GPS2Vec}.  This application consists
 of a neural network that would generate embeddings for a location, based on the semantic 
@@
 With the integration of these embeddings, a successful geo-tagged image classification 
-task is demonstrated. In another effort  \parencite{PlaceDeduplication}, the location 
-embeddings derived from user content such as addresses, phone numbers, images, place
+task is demonstrated. 
+\par In another effort\parencite{PlaceDeduplication}, the location embeddings derived from user content such as addresses, phone numbers, images, place
 names and all other details from multiple individual online posts, are compressed to 
@@
 location embeddings, the application can identify the duplicates and eliminate them to 
-result in an efficient place graph.\par
-In 2018, a new approach, “DeepMove” is proposed \parencite{DeepMove}. This 
+result in an efficient place graph.
+\par In 2018, a new approach, “DeepMove” is proposed \parencite{DeepMove}. This 
 implementation employs Trip model and OD model to learn latent representations of 
@@
 implementation had been demonstrated that would derive location embeddings from human 
-trajectories \parencite{DisentangledLocationEmbeddings}. With the ever- growing 
+trajectories \parencite{DisentangledLocationEmbeddings}. 
+\par With the ever-growing 
 footprint of digital and social media platforms in everyday life of an individual, a
@@
 different locations. LeGo-CM is one such effort that would generate embeddings of 
-geotagged tweets \parencite{GeotaggedTweets}. These resultant embeddings could be 
-applied to understand the kind of content each location reflects.Venue2Vec is another 
+geotagged tweets \parencite{GeotaggedTweets}. The resulting embeddings could be 
+applied to understand the kind of content each location reflects. 
+\par Venue2Vec is another 
 idea that is implemented to enhance the location-aware services in various platforms 
@@
 based on the input features. Hence, the model is effective in predicting the next and  
-future check-in location. Community Flow prediction is another area of interest where
+future check-in location. 
+\par Community Flow prediction is another area of interest where
 embeddings are a major problem solver. Geo-contextual Multitask Embedding Learner (GMEL)
@@
 
-We used two different methods to reduce the dimensionality of our data - Principle Component
-Analysis and undercomplete autoencoders. A simple linear regression was performed on 
-the encoded data that was generated by both the PCA and AE models. Additionally, we performed
-a linear regression on the unencoded values for comparison. Each model's goodness of fit
-was evaluated using the percentage of explained variance for the three linear regressions.
+\par Two different methods were used to reduce the dimensionality of the data - principle component analysis (PCA) and autoencoders.
+A simple linear regression was performed on the encoded data that was generated by both the PCA and autoencoder models.
+Additionally, a linear regression on the unencoded values was performed for comparison.
+Each model's goodness of fit was evaluated using the percentage of explained variance of the three linear regressions.
 
-% pipeline
-\begin{figure}[h]
-\includegraphics[width=\textwidth, height=6cm]{/home/nicks/Documents/paper-draft/images/pipeline.png}
+\par In order to explore the nature of the dataset, we calculated the Pearson correlation coefficient of the normalized timeseries data.
+The Pearson correlation coefficient allows us allows us to observe the strength of the linear correlation between features.
+
+\begin{equation*}
+    r_{XY}=\frac{cov(X,Y)}{\sigma_{X} \cdot \sigma_{Y}}
+\end{equation*}
+
+% Correlation Matrix 
+\begin{figure}[h!]
+\includegraphics[width=\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/corr_matrix.png}
+\caption{Correlation Matrix}
+\label{fig:matrix}
+\end{figure}
+
+\par As previously mentioned in the introduction, the correlation between features was revealed to be rather high.
+In particular, days one hundred thirty-five to one hundred-eighty nearly all have a correlation coefficient greater than 0.75.
+This resulted in some overfitting which can be seen in the autoencoder and pca explainable variance scores, which were greater than the unencoded scores.
+
+\newpage
+\begin{Center}
+\begin{figure}[h!]
 \caption{Data Pipeline}
 \label{fig:pipeline}
+    \hspace{1.5cm}\includegraphics[width=10cm, height=13cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pipeline.png}
 \end{figure}
+\end{Center}
 
@@
 eight polluting gases and particulates: CO, NO, NO$_2$, O$_3$, SO$_2$, NH$_3$, PM$_{2.5}$, 
-and PM$_{10}$. The dataset also includes the hourly timestamp, the latitude and longitude (part of the input),
-and the air quality index, the latter of which we did not end up using.\par
-We looked at data for the eight pollutants over a six-and-a-half month time 
-frame: January 27th, 2020 through June 6th, 2021. Our final 
-dataset included just over eighteen thousand cities in the United States and other U.S. 
-territories. We found it easier to focus on a single gas when developing 
-the models, so we chose to concentrate on CO during the engineering process. Given
-enough time and computational power, the techniques performed on the CO model would be 
-easily extensible and transferable for the other seven pollutants.
+and PM$_{10}$.
+Also included was the city name, hourly timestamp, the latitude and longitude (part of the input), and the air quality index (aqi).
+We chose not to use the aqi because it is overall less sensitive to change and is unlikely to reveal any long-term patterns over such a short time span .
+\par We looked at data for the eight pollutants over a five month time frame: January 27th, 2020 through June 6th, 2021.
+The final dataset includes just over eighteen thousand cities in the United States and other U.S.
+territories.
+We found it easier to focus on a single gas when developing the models, so we chose to concentrate on CO during the engineering process.
+The same process was then applied to the other seven gases and particulates after the development was completed.
 
@@
         Time interval & Daily  \\
-        $\#$ of cities & 18,526 \\
-        $\#$ of features & 191 \\
-        $\#$ of component gases & 8 \\
-        Dependent variable & Single day of time series (06/05/21) \\
+        Cities & 18,526 \\
+        Features & 190 \\
+        Component gases & 8 \\
+        Dependent variable & Last day of the period (06/05/21) \\
         \hline
@@
 \end{table}
+\par OpenWeather's Pollution API ordinarily returns hourly values, but we opted to use the daily average pollution levels as our feature set in order to make the number of starting dimensons more manageable.
+The daily averages were also feature engineered to be proportional to the number of features so as to avoid the curse of dimensionality \parencite{Trunk79}.
+\par The original data was unit normalized so that the gradient descents could converge more quickly while still preserving the original differences in range.
+Each step was saved separately to provide expeditious access to both the normalized and non-normalized data.
 
-OpenWeather's Pollution API ordinarily returns hourly values, but we opted to use the daily
-average pollution levels as our feature set in order to make the number of starting dimensons
-more manageable. We resolved to feature engineer the daily 
-averages and proportional to the number of features and thus avoid the curse of dimensionality. 
-\parencite{Trunk79}. \note{Idk about this} \par
-We separated the dependent from the input before unit normalizing the input. 
-The normalized data was then encoded by each model.
+\subsection{Autoencoder Model}
+\par The autoencoder model used consists of a single encoded and decoded layer.
+The data was split into a train and test set at the beginning of the testing cycle, and was further separated into a validation set for each given dimension.
+The validation set was essential for updating the model weights after every epoch.
+After the model was trained and validated, we created the finalized embeddings using the full dataset.
+We created embeddings for several of the most significant dimensions based off their increase in explained variance scores.
 
-\subsection{Undercomplete Autoencoder Model}
-The autoencoder model used consists of an encoded and decoded layer. We performed
-a grid search for the CO model to determine the optimal hyperparameters. The following 
-hyperparameters were tested:
+\subsubsection{Model Validation}
+\par A grid search was performed for each model to determine the optimal hyperparameters. The following hyperparameters were tested:
 \begin{itemize}
@@
 \end{itemize}
-K-folds cross validation was used to validate the grid search using five folds. We selected the
-highest coefficient of determination that appeared in each of the five folds for every given
-dimension. The most frequently occuring combination of hyperparameters was then used to create
-the finalized embedding.
+\par K-folds cross validation was implemented to validate the grid search using five folds.
+Additional, the training split for each dimension was separated into a validation set.
+Each hyperparameter was selected based on the highest coefficient of determination that appeared in each of the five folds for every given dimension.
+The most frequently occuring combination of hyperparameters was then used to create the finalized embedding.
+The lowest coefficients of determination were saved and compared with the highest values to give an upper and lower bound of expected values.
+\par After some preliminary testing, it was decided that the activation, optimizer, and loss functions should remain constant.
+The exponential increase in computing time this would have on the grid search drastically outweighed any minor performance gains that a thorough optimization of these functions may have yielded.
 
 \begin{table}[h!]
-    \caption{Dataset Characteristics}
+    \caption{Constant hyperparameters used across all autoencoder models}
     \label{tab:table1}
@@
         \hline
-        \multicolumn{2}{c}{Optimal Hyperparameters for Carbon Monoxide (CO)} \\
+        \multicolumn{2}{c}{Constant Hyperparameters}\\
         \hline
@@
         Optimizer & Adam \\
-        Learning rate & 0.001 \\
-        Batch size & 64 \\
-        Epochs & 150 \\ 
         \hline
@@
 \end{table}
+Moreover, grid search was only performed on a select number of key dimensions rather than on all one hundred ninety dimensions.
+The first ten dimensions, as well as every tenth dimension until one hundred-twenty were tuned.
+For each intermediary dimension that did not have a set of key hyperparameters, the hyperparameters of the previous key dimension was used.
+Once again, the tradeoff between minor performance gains and large compute times led us to adjust the scope of the  hyperparameter tuning.
+
+\begin{table}[h!]
+    \caption{Grid Search Results for Carbon Monoxide}
+    \label{tab:table2}
+    \vspace{0.1cm}
+    \begin{tabular}{p{2cm}p{2cm}p{2cm}p{2cm}p{2cm}}
+        \hline
+        \multicolumn{5}{c}{CO Grid Search Key Dimensions 1-10}\\
+        \hline
+        dimension & $r^{2}$ & learning rate & batch & epochs  \\
+        1 & 0.22 & 0.01 & 64 & 10 \\
+        2 & 0.44 & 0.0001 & 256 & 50\\
+        3 & 0.49 & 0.001 & 64 & 50\\
+        4 & 0.52 & 0.0001 & 32 & 50\\
+        5 & 0.56 & 0.0001 & 32 & 100\\
+        6 & 0.57 & 0.01 & 32 & 10\\
+        7 & 0.60 & 0.0001 & 32 & 150\\
+        8 & 0.63 & 0.0001 & 64 & 150\\
+        9 & 0.62 & 0.0001 & 32 & 150\\
+        10 & 0.62 & 0.0001 & 64 & 100\\
+        \hline
+    \end{tabular}
+\end{table}
+\begin{table}[h!]
+    \caption{Grid Search Results for Carbon Monoxide}
+    \label{tab:table2}
+    \vspace{0.1cm}
+    \begin{tabular}{p{2cm}p{2cm}p{2cm}p{2cm}p{2cm}}
+        \hline
+        \multicolumn{5}{c}{CO Grid Search Key Dimensions 20-120}\\
+        \hline
+        dimension & $r^{2}$ & learning rate & batch & epochs  \\
+        20 & 0.62 & 0.01 & 64 & 50\\
+        30 & 0.64 & 0.01 & 128 & 75\\
+        40 & 0.67 & 0.01 & 256 & 100\\
+        50 & 0.68 & 0.01 & 64 & 75\\
+        60 & 0.69 & 0.001 & 32 & 150\\
+        70 & 0.70 & 0.001 & 64 & 150\\
+        80 & 0.71 & 0.001 & 32 & 150\\
+        90 & 0.71 & 0.01 & 64 & 100\\
+        100 & 0.72 & 0.01 & 64 & 150\\
+        110 & 0.72 & 0.01 & 32 & 75\\
+        120 & 0.73 & 0.001 & 64 & 100\\
+        \hline
+    \end{tabular}
+\end{table}
+\newpage
 
 \subsection{Principle Component Analysis}
-The PCA model was trained with k-fold cross-validation using five folds. The dataset was
-split into a training, validation, and hold-out test set.
+The PCA model was trained with k-fold cross-validation using five folds.
+The dataset was split into a training and test set.
+The trained PCA models were used to create embeddings at different scales of dimensionality based on the number of input components.
+We prioritized the first ten dimensions because they generally represent the largest increase in the percent explained variance.
 
-\subsection{Model Validation}
-K-folds cross validation was performed on the linear regression test using five folds. The folds
-with the best and worst variance scores were extracted so as to make an equitable evaluation of
-the model.
+\subsection{Additional Model Validation}
+In addition to the PCA and Autoencoder models, a linear regression was performed on the unencoded features.
+The resulting explainable variance scores serve as a benchmark with which we can compare each model's data loss.
+
+\begin{figure}[h!]
+    \centering
+    \caption{Unencoded R2 Scores}
+    \label{fig:pipeline}
+    \includegraphics[width=0.8\linewidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/unencoded_r2.png}
+\end{figure}
 
 \section{Results}
-We compared the percent explained variance across one hundred twenty dimensions for
-the encoded values generated by the PCA and autoencoder models as well as for the unencoded
-values by using a simple linear regression. While PCA was overall less noisy, we found that
-the autoencoder model achieved a much higher explainable variance for the first few dimensions.
+We compared the percent explained variance across one hundred ninety dimensions for the encoded values generated by the PCA and autoencoder models.
+While PCA was overall less noisy, we found that the autoencoder model achieved a much higher explainable variance for the first several dimensions.
+\comment{
 \\ \note{Note that results graphs are placeholders}
 % ae vs pca linegraph
+
 \begin{figure}[h!]
-
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nicks/Documents/paper-draft/images/ae_all_dims.png} 
-    \caption{Autoencoder Dimensionality Reduction}
+    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
+    \caption{PCA Dimensionality Reduction}
     \label{fig:ae_dim_reduction}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nicks/Documents/paper-draft/images/ae_all_dims.png}
+    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/ae_all_dims.png}
     \caption{PCA Dimensionality Reduction}
@@
 \end{subfigure}
-
 \caption{Graph of autoencoder and PCA reduced representations of air pollutants}
 \label{fig:ae_vs_pca}
+\end{figure}
+}
+
+\begin{figure}[h!]
+    \includegraphics[width=\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
+    \caption{PCA Dimensionality Reduction}
+    \label{fig:ae_dim_reduction}
+\end{figure}
+
+\begin{figure}[h!]
+    \includegraphics[width=\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
+    \caption{Autoencoder Dimensionality Reduction\ \note{Placeholder for AE}}
+
+    \label{fig:ae_dim_reduction}
 \end{figure}
@@
 \begin{figure}[h!]
-
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nicks/Documents/paper-draft/images/co_scatter_ae_wlist.png} 
+    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/co_scatter_ae_wlist.png} 
     \caption{Outlier Cities}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nicks/Documents/paper-draft/images/co_scatter_ae_large_cities.png}
+    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/co_scatter_ae_large_cities.png}
     \caption{Densely Populated Cities}
@@
 
-By selecting from a list the top two hundred highly-populated cities, we found that nearly
-all of them were located near the center of the clustering. Conversely, we found that many 
-of the outliers were smaller cities in rural locations. \par
+\par By selecting from a list the top two hundred most highly-populated cities in the US, we found that nearly all of them were located near the center of the clustering.
+Conversely, we found that many of the outliers were smaller cities in rural locations.
 
-\newpage
-
-To further explore the meaning of the hidden dimensions, we calculated the Pearson correlation
-coefficient of the normalized timeseries data and the embeddings. The Pearson correlation
-coefficient allows us allows us to observe the strength of the linear correlation between
-the encoded and unencoded values. 
-\begin{equation*}
-    r_{XY}=\frac{cov(X,Y)}{\sigma_{X} \cdot \sigma_{Y}}
-\end{equation*}
-\textit{Where cov(X,Y) is the covariance between X and Y \\
-and $\sigma_{X},\sigma{Y}$ are the standard deviations of X and Y respectively}
-
-\vspace{0.1in}
-
-We calculated the correlation coefficients and created a correlation matrix of the days that had
-correlation coefficients of the highest magnitude.\par
-
-% pipeline
-\begin{figure}[h]
-\includegraphics[width=\textwidth]{/home/nicks/Documents/paper-draft/images/corr_matrix_25.png}
-\caption{Correlation Matrix}
-\label{fig:matrix}
-\end{figure}
 \section{Discussion}

### 2021-12-21 src/autoencoder.py

@@ -74,7 +74,7 @@ def ae_train_test(dims, X_train, X_test, Y_train, Y_test, component, param_grid)
         param_grid: List of key hyperparameters to test 
     '''
 
-    # Metrics file
+    # Metrics file to write to
     file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/ae/{component}_metrics'
     # Train Autoencoder model
     variance_list = []
@@ -164,12 +164,13 @@ def ae_run(dim, X, X_train, Y_train, lr, batch, epochs, component, cities):
     vector_data.to_csv(path_or_buf=vec_file, index=None)
 
 
-def grid_search(x, y, folds, component, iter_dims, key_params):
+def grid_search(file_name, x, y, folds, component, iter_dims, key_params):
     '''
     Run the linear regression for a single gas
     using k-fold cross validation strategy
 
     @params:
+        file_name: The output file two write the gs results to
         x: The normalized x values to be used in the embedding
         y: The dependent variable
         folds: Number of folds to test
@@ -179,7 +180,6 @@ def grid_search(x, y, folds, component, iter_dims, key_params):
         batch: Batch size
         epochs: Number of epochs        
         iter_dims: Key dimensions to perform the grid search on
-        key_params: list of key parameters to test
     '''
 
     print(f'---------- Beginning Linear Regression for {component} ----------')
@@ -187,13 +187,11 @@ def grid_search(x, y, folds, component, iter_dims, key_params):
     # k-fold cross validation
     kfold = KFold(n_splits=folds, shuffle=True, random_state=10)
     fold_count=0
-    # File name
-    file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/grid_params/{component}_vec.csv'
     
     # Headers for grid search
     grid_list = ['fold', 'dim', 'variance', 'r2', 'lr', 'batch', 'epochs']
     # Write header
-    with open(file_name,'w', newline='') as f:
+    with open(file_name,'w+', newline='') as f:
         writer = csv.writer(f)
         writer.writerow(grid_list)
         f.close() 
@@ -216,9 +214,9 @@ def grid_search(x, y, folds, component, iter_dims, key_params):
                     Y_train=Y_train,
                     component=component,
                     activation=('tanh', 'tanh'),
[…]

### 2021-12-21 src/distributed_gs.py

@@ -1,0 +1,70 @@
+import autoencoder as ae
+import numpy as np
+import itertools
+import pandas as pd
+import os
+import sys
+from sklearn.model_selection import train_test_split
+
+
+def main(*argv):
+    # Starting dimensions
+    dims = 190
+    # K-fold folds
+    folds = 5
+    # Grid search params
+    lr = [0.0001, 0.001, 0.01, 0.1]
+    batch = [32, 64, 128, 256]
+    epochs = [10, 50, 75, 100, 150]
+    # Param vector
+    iter_dims = []
+    # List of key params to test
+    key_params = list(itertools.product(lr,batch,epochs))
+    ### Input files ###
+    # Open normalized data and dependent, non-normalized data
+    dfx = pd.read_csv(f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/data/data_norm/co_data_norm.csv")
+    dfy = pd.read_csv(f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/data/data_clean/co_data_clean.csv")
+    # City names to append
+    cities = dfy['city'].values
+    # Grid params to create model with
+    param_grid = pd.read_csv(f'/home/nick/github_repos/Pollution-Autoencoders/data/grid_params/hyperparams.csv')
+    # Name of the component gas
+    component = 'no2'
+    file_name=''
+    # Set x as the normalized values, y (non-normalized) as the daily average of final day
+    X = dfx.values
+    Y = dfy.loc[:, ['co_2021_06_06']].values
+    # Split into train/test data
+    X_train, X_test, Y_train, Y_test = train_test_split(X, Y, test_size=0.20, random_state=40)
+    
+    # Arguments
+    for args in sys.argv:   
+        if args == "-k_1_5" or args == "--first-five":
+            iter_dims = np.arange(1,6,1)
+            file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/grid_params/{component}_grid_params_1_5.csv'
+        elif args == "-k_6_9" or args == "--six-to-nine":
+            iter_dims = np.arange(6,10,1)
+            file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/grid_params/{component}_grid_params_6_9.csv'
+        elif args == "-k_10_60" or args == "--ten-to-fifty":
+            iter_dims = np.arange(10,51,10)
[…]

### 2021-12-21 src/pca.py

@@ -84,17 +84,21 @@ def pca_train_test(dims, x, y, folds, component):
                 f.close() 
 
 
-# TODO: FIX THIS MONSTROSITY!!!!!
-def interpolate(component):
+def interpolate(dims, component):
     ''' Function to derive the best and worst fold for a given dimension '''
 
-    df = pd.read_csv(f'/home/nick/github_repos/Pollution-Autoencoders/data/test_metrics/pca/{component}_test_metrics.csv')
+    df = pd.read_csv(f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/pca/{component}_metrics')
     # Outputs
-    best_metrics = f'/home/nick/github_repos/Pollution-Autoencoders/data/test_metrics/pca/derived/{component}_best_metrics.csv'
-    worst_metrics = f'/home/nick/github_repos/Pollution-Autoencoders/data/test_metrics/pca/derived/{component}_worst_metrics.csv'
+    best_metrics = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/pca/best_worst/{component}_best_metrics.csv'
+    worst_metrics = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/pca/best_worst/{component}_worst_metrics.csv'
     # Lists to write to file
     best_list=worst_list=['dim', 'variance', 'r2']
 
+    # Debug
+    #print('dim 1 ', df['variance'][0])
+    #print('dim 190 ', df['variance'][189])
+
+    
     # Write headers for best and worst metrics
     with open(best_metrics,'w', newline='') as f:
         writer = csv.writer(f)
@@ -104,15 +108,18 @@ def interpolate(component):
         writer = csv.writer(f)
         writer.writerow(worst_list)
         f.close() 
+    
 
-    for i in range(189):
+    # Fold range that describes indexes of all dimensions
+   
+    for i in range(dims):
         # Reset best and worst variance and index for every dim
         best_var, worst_var = 0, 100
         best_idx=worst_idx=0
 
         # Retrieve a dict of each fold's variance for a given dimension
-        fold_var = { i : df['variance'][i], i+189 : df['variance'][i+189], i+378 : df['variance'][i+378], 
-            i+567 : df['variance'][i+567], i+756 : df['variance'][i+756] }
[…]

### 2021-12-21 src/visualization.py

@@ -5,7 +5,7 @@ import seaborn as sns
 import itertools
 
 
-def linegraph(f_name, type, dims, component_names, colors_list):
+def linegraph(type, dims, component_names, colors_list):
     ''' 
     Plot the explained variance across dimensions using the model results
 
@@ -18,17 +18,17 @@ def linegraph(f_name, type, dims, component_names, colors_list):
     '''
     
     num_of_comp = list(range(2,dims+1)) # TODO: delete this later
-    num_of_dims = list(range(1, dims+1))
+    num_of_dims = list(range(2, dims+1))
     # Plot results
     plt.figure(figsize=(12,10))
     for i, component in enumerate(component_names):
         # AE or PCA
         if type == 'ae':
             #file_name = f'{f_name}{component}_metrics.csv'
-            file_name = '/home/nicks/github_repos/Pollution-Autoencoders/data/model_metrics/test_metrics.csv'
+            file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/ae/{component}_metrics.csv'
             plt_title = 'Autoencoder Reduced Representation of Air Pollutants'
         elif type == 'pca':
-            file_name = f'{f_name}{component}_metrics.csv'
+            file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/pca/{component}_metrics'
             plt_title = 'PCA Reduced Representation of Air Pollutants'
         else:
             print('Type must be "ae" or "pca"')
@@ -39,15 +39,25 @@ def linegraph(f_name, type, dims, component_names, colors_list):
         variance = model['variance']
         r2 = model['r2']
         
-
-        plt.plot(num_of_comp, variance[:dims-1], label = '{}'.format(component), linestyle = '-', marker = '+', color = colors_list[i])
-        plt.plot(num_of_comp, r2[:dims-1], linestyle = '-.', marker = 'H', color = colors_list[i])
+        # sns
+        ax = sns.lineplot(x=num_of_dims, y=variance[:dims-1], linewidth=1, label=f'{component}', color=colors_list[i])
[…]

### 2021-12-21 test.py

@@ -1,0 +1,10 @@
+
+x=2437
+y=875
+
+while (x!=y):
+    if (x>y):
+        x=x-y
+    elif (x<y):
+        y=y-x
+    print(f'x: {x}, y: {y}')

### 2021-12-24 paper-draft/paper.tex
Added 3D scatterplot; reverse geolocater to add state and country labels

@@
 
-\subsubsection{Model Validation}
+\begin{table}[h!]
+    \caption{Model Splits}
+    \label{tab:table1}
+    \vspace{0.1cm}
+    \begin{tabular}{p{4cm}p{7cm}}
+        \hline
+        \multicolumn{2}{c}{Constant Hyperparameters}\\
+        \hline
+        Training Size & 14,080\\
+        Testing Size & 3,705\\
+        Validation Size & 741\\
+        Total Size & 18,526\\
+        \hline
+    \end{tabular}
+\end{table}
 \par A grid search was performed for each model to determine the optimal hyperparameters. The following hyperparameters were tested:
@@
 
+Although the model provides insights on the most succinct features that characterize urban air pollution, more specific external data is needed to uncover these features.
+One study \parencite{Bai2019} uses a stacked autoencoder approach and focuses on the seasonality of pm2.5 levels.
+The proposed model considers both the seasonality and temporality of pm2.5 levels and supporting meteorological data.
+The research group was able to accurately predict the pm2.5 levels to the hour for a given season.
+
 \section{Conclusion}

### 2021-12-24 src/autoencoder.py

@@ -124,7 +124,7 @@ def ae_train_test(dims, X_train, X_test, Y_train, Y_test, component, param_grid)
     metrics_data.to_csv(path_or_buf=file_name, index=False)
 
 
-def ae_run(dim, X, X_train, Y_train, lr, batch, epochs, component, cities):
+def ae_run(dim, X, X_train, Y_train, optimal_hyperparams, component, cities):
     '''
     Run the autoencoder model by performing a cross-validated linear
     regression on the encoded data
@@ -149,9 +149,9 @@ def ae_run(dim, X, X_train, Y_train, lr, batch, epochs, component, cities):
         Y_train=Y_train,
         component=component,
         activation=('tanh', 'tanh'),
-        lr=lr, 
-        batch=batch,
-        epochs=epochs
+        lr=optimal_hyperparams['lr'][dim], 
+        batch=optimal_hyperparams['batch'][dim],
+        epochs=optimal_hyperparams['epochs'][dim]
     )
    
     # Create encoded data based off full dataset
@@ -245,7 +245,7 @@ def main():
     #component_names = ['co', 'no', 'no2', 'o3', 'so2', 'pm2_5', 'pm10', 'nh3']
     component_test = 'o3'
     # Starting dimensions
-    dims = 190
+    dims = 2
     # K-fold folds
     folds = 5
     # Grid search params
@@ -256,8 +256,8 @@ def main():
     key_params = list(itertools.product(lr,batch,epochs))
     # List of key dimensions to perform grid search on
     #iter_dims = np.append(np.arange(1, 10, 1), np.arange(10, 120, 10))
-    iter_dims = np.arange(1,10,1)
-    #iter_dims = np.arange(10,121,10)
+    #iter_dims = np.arange(1,10,1)
+    iter_dims = np.arange(10,121,10)
 
     ### Input files ###
     # Open normalized data and dependent, non-normalized data
@@ -275,15 +275,16 @@ def main():
     ### Function calls ###
 
     ## Grid Search
-    output_file = f'/home/nick/github_repos/Pollution-Autoencoders/data/grid_params/{component_test}_grid_params_1_9.csv'
-    grid_search(output_file, X, Y, folds, component_test, iter_dims, key_params)
+    #output_file = f'/home/nick/github_repos/Pollution-Autoencoders/data/grid_params/{component_test}_grid_params_10_120.csv'
[…]

### 2021-12-24 src/reverse_geocode.py

@@ -1,0 +1,54 @@
+import pandas as pd
+import googlemaps
+import urllib.request
+import json
+#import dotenv
+
+# Reverse GeoCoding to populate data with state/country codes
+# 40.714224, -73.961452
+def rev_geo(lat, lon, key):
+    REVGEO_BASE_URL = 'https://maps.googleapis.com/maps/api/geocode/json'
+    URL_PARAMS = f'latlng={lat},{lon}&key={key}'
+    
+    url = REVGEO_BASE_URL + "?" + URL_PARAMS
+    #print(f'url is {url}')
+    # Read the contents of the generated url and decode the result
+    with urllib.request.urlopen(url) as f:
+        response = json.loads(f.read().decode())
+    status = response["status"]
+    if status == "OK":
+        result = response["results"][0]['address_components']
+        #print(result)
+
+        state_code = country_code = 'NULL'
+        # Parse the response object for administrative_area_level_1 and country types
+        for r in result:
+            if r['types'][0] == 'administrative_area_level_1':
+                state_code = r['short_name']
+            if r['types'][0] == 'country':
+                country_code = r['short_name']
+    else:
+        print(status)
+        print(response['error_message'])
+        rev_geo_data = None
+    return (state_code, country_code)
+
+KEY = 'AIzaSyCZr4o4NiY5tKHlWNVTh9ExeiKQ_O35UMc'
+#LAT = 30.2671500
+#LON = -97.7430600
+component_names = ['co', 'no', 'no2', 'o3', 'so2', 'pm2_5', 'pm10', 'nh3']
+states = countries = []
+
+for component in component_names:
+    df = pd.read_csv(f'./data/data_clean/{component}_data_clean.csv')
+    for lat, lon in zip(df['lat'], df['lon']):
+        #print(f'{lat} , {lon}')
+        state_code, country_code = rev_geo(lat, lon, KEY)
+        states.append(state_code)
+        countries.append(country_code)
+        
+    # Add new columns to the df
+    df.insert(loc = 1, column = 'state', value = states)
+    df.insert(loc = 2, column = 'country', value = countries)
+    df.to_csv(f'/home/nick/Downloads/revised_data/{component}_data_clean.csv')
+#print(data)

### 2021-12-24 src/visualization.py

@@ -3,6 +3,8 @@ import numpy as np
 import matplotlib.pyplot as plt
 import seaborn as sns
 import itertools
+import pyparsing
+from mpl_toolkits import mplot3d
 
 
 def linegraph(type, dims, component_names, colors_list):
@@ -117,56 +119,6 @@ def metrics_comparison(type, dims, component_names, colors_list):
         plt.legend()
     plt.show()
 
-def scatter(f_name, type, component, color):
-    '''
-    Autoencoder scatter plot of multiple components for the first two dimensions
-
-    @params:
-        f_name: Name of the file
-        component: Name of gas or particulate
-        color: Scatter plot color
-    '''
-
-    # Read in X_train values for first two dimensions from model results
-    if type == 'ae':
-        file_name = '%s/autoencoders/trail1/%s_ae_results_gs.csv' % (f_name, component)
-    elif type == 'pca':
-        file_name = '%s/autoencoders/trial1/%s_pca_results_gs.csv' % (f_name, component)
-    else:
-        print('Type must be "ae" or "pca"')
-        quit()
-
-    model = pd.read_csv(filepath_or_buffer=file_name)
-    X_train = pd.DataFrame(data=model)
-    X = X_train['{}_X_train_ax1'.format(component)]
-    Y = X_train['{}_X_train_ax2'.format(component)]
-
-    # Read in city data frame and select list of cities
-    city_df = pd.read_csv(filepath_or_buffer='/home/nicks/github_repos/Pollution-Autoencoders/data/data_clean/{}_data_clean.csv'.format(component))
-    annotations = pd.DataFrame(data=city_df)
-   
-    # Read in whitelisted city data
-    # Whitelist can be top200 cities or the outliers
-    wlist_data = pd.read_csv(filepath_or_buffer='/home/nicks/github_repos/Pollution-Autoencoders/data/other/top200.csv')
-    wlist = pd.DataFrame(data=wlist_data[:20])
-    
-    # Plot figure
-    plt.figure(figsize=(14,14))
-    plt.rcParams.update({'font.size': 18})
-    plt.scatter(X, Y, label='{}'.format(component), c=color, alpha=0.1)
-
-    size = len(X_train)
[…]

### 2021-12-24 src/reverse_geocode.py
maintenance

@@ -2,6 +2,7 @@ import pandas as pd
 import googlemaps
 import urllib.request
 import json
+import keys
 #import dotenv
 
 # Reverse GeoCoding to populate data with state/country codes
@@ -33,7 +34,7 @@ def rev_geo(lat, lon, key):
         rev_geo_data = None
     return (state_code, country_code)
 
-KEY = 'AIzaSyCZr4o4NiY5tKHlWNVTh9ExeiKQ_O35UMc'
+KEY = key.GMAPS_KEY
 #LAT = 30.2671500
 #LON = -97.7430600
 component_names = ['co', 'no', 'no2', 'o3', 'so2', 'pm2_5', 'pm10', 'nh3']

### 2022-02-08 paper-draft/paper.tex
Fixed autoencoder training function; updated paper

@@
 \usepackage[document]{ragged2e}
-\usepackage[pass, letterpaper]{geometry}
+%\usepackage[pass, letterpaper]{geometry}
+\usepackage[margin=0.75in]{geometry}
 \usepackage{amsmath}
@@
 \end{abstract}
+
 \vspace{0.25cm}
@@
 
-\par Over the past several decades, air pollution has been a major concern for human health
+\par For the past several decades, air pollution has become a major concern for human health
 and the environment. The EPA sets national air quality standards for  six major air
@@
 \par The idea of generating embeddings for a location is a highly potential expanse for
-research. The two major forces behind its competency are, the fact that large volumes of
+research. The two major forces behind its competency are the fact that large volumes of
 data is being effectively compressed and that these embeddings, when plugged into any
@@
 \begin{figure}[h!]
-\includegraphics[width=\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/corr_matrix.png}
-\caption{Correlation Matrix}
+\begin{center}
+\includegraphics[width=0.7\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/corr_matrix.png}
+\end{center}
+    \caption{Correlation Matrix for Carbon Monoxide (daily averages for cities throughout the US)}
 \label{fig:matrix}
@@
 \label{fig:pipeline}
-    \hspace{1.5cm}\includegraphics[width=10cm, height=13cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pipeline.png}
+    \begin{center}
+    \includegraphics[width=10cm, height=13cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pipeline.png}
+    \end{center}
 \end{figure}
@@
 \begin{table}[h!]
+\begin{center}
     \caption{Dataset Characteristics}
@@
     \end{tabular}
+\end{center}
 \end{table}
@@
 \begin{table}[h!]
+\begin{center}
     \caption{Model Splits}
@@
 \begin{table}[h!]
+\begin{center}
     \caption{Constant hyperparameters used across all autoencoder models}
@@
 \begin{table}[h!]
+\begin{center}
     \caption{Grid Search Results for Carbon Monoxide}
@@
     \end{tabular}
+\end{center}
 \end{table}
 \begin{table}[h!]
+\begin{center}
     \caption{Grid Search Results for Carbon Monoxide}
@@
 
+\newpage
+
 \section{Results}
-We compared the percent explained variance across one hundred ninety dimensions for the encoded values generated by the PCA and autoencoder models.
+
+\par We compared the percent explained variance across all one hundred ninety dimensions for the encoded values generated by the PCA and AE models.
 While PCA was overall less noisy, we found that the autoencoder model achieved a much higher explainable variance for the first several dimensions.
-\comment{
-\\ \note{Note that results graphs are placeholders}
-% ae vs pca linegraph
 
+\begin{center}
 \begin{figure}[h!]
-\begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
-    \caption{PCA Dimensionality Reduction}
-    \label{fig:ae_dim_reduction}
-\end{subfigure}% <- % sign must be here!
-\begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.1\linewidth, height=4cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/ae_all_dims.png}
+    \centering
+    \includegraphics[width=0.7\linewidth, height=7cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
     \caption{PCA Dimensionality Reduction}
     \label{fig:pca_dim_reduction}
-\end{subfigure}
-\caption{Graph of autoencoder and PCA reduced representations of air pollutants}
-\label{fig:ae_vs_pca}
-\end{figure}
-}
-
-\begin{figure}[h!]
-    \includegraphics[width=\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
-    \caption{PCA Dimensionality Reduction}
-    \label{fig:ae_dim_reduction}
 \end{figure}
@@
 \begin{figure}[h!]
-    \includegraphics[width=\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
+    \centering
+    \includegraphics[width=0.7\linewidth, height=7cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
     \caption{Autoencoder Dimensionality Reduction\ \note{Placeholder for AE}}
-
     \label{fig:ae_dim_reduction}
 \end{figure}
+\end{center}
 
-\par We clustered the cities based on the first two latent dimensions in order to see how
-the model characterizes each city
+\par In nearly every case shown in figures \ref{fig:pca_dim_reduction} and \ref{fig:ae_dim_reduction}, the autoencoder model shows consistently higher explainability than the PCA.
+Tropospheric Ozone (O3) in particular shows a massive jump of over fifty percentage points explainable variance in the first two dimensions.
+For this reason, O3 was the primary model for the subsequent visualizations.
+\par In order to better visualize how the hidden layers distill information, we clustered the cities based on the first two encoded dimensions.
+Additionally, a number of location heuristics were applied to further understand the distribution of cities.
+Figure \ref{fig:heuristic_labels} shows how the embeddings are clustered by regions, as well as by individual states and territories.
 
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/co_scatter_ae_wlist.png} 
-    \caption{Outlier Cities}
+    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/region.png} 
+    \caption{Model characterization grouped by region}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/outside-us.png}
+    \caption{Model characterization of U.S. Territories}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/california.png} 
+    \caption{Model characterization of California}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/texas.png} 
+    \caption{Model characterization of Texas}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/new-york.png}
+    \caption{Model characterization of New York}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/minnesota.png}
+    \caption{Model characterization of Minnesota}
+\end{subfigure}
+\caption{Scatter plots of outlier and highly-populated cities over first two dimensions}
+\label{fig:heuristic_labels}
+\end{figure}
+
+\par The model's characterization appears to be heavily dependent on geographic location.
+When grouped by region, one can see that there is a fairly tight distribution, especially among the Midwest and Northeast regions.
+Puerto Rico is also tightly clustered, but the territory contains a limited 211 representative data points.
+
+\par Curiously enough, population size does not seem to be one of the most succint features of the model.
+Figure \ref{fig:population_comparison} shows the top two hundred most populous cities in the US for the first four dimensions of the model.
+The first four latent dimensions of O3 account for nearly 60\% of the total explainable variance.
+However, the top 200 cities appear to be randomly dispersed throughout, suggesting that population alone has less of an effect on the model.
+It is possible that population is represented indirectly
+
+\begin{figure}[h!]
+\begin{subfigure}{0.5\textwidth}
+    \includegraphics[width=1.0\linewidth, height=10cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/population.png} 
+    \caption{Model characterization grouped by region}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+    \includegraphics[width=1.0\linewidth, height=10cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/population_3and4.png}
+    \caption{Model characterization of U.S. Territories}
+\end{subfigure}
+\caption{Scatter plots of outlier and highly-populated cities over first two dimensions}
+\label{fig:population_comparison}
+\end{figure}
+
+\par It is difficult to determine why the model does not seem to consider population size.
+One reason for this may be that temperature and climate "overpowers" any influence a city's size may have.
+Any number of secondary effects related to climate and location could pull the model away from considering population.
+Moreover, The time frame of our data spans from winter 2020 to early summer of 2021, which is not long enough to fully capture seasonality.
+Given a larger data set, it is likely that both population and seasonality would play much larger roles.
+
+\par Due to the importance of geography in the embedding, we created several geographic scatter plots to visualize both the embedding values and pollution levels with respect to location.
+
+% Scattergeo 
+\begin{figure}[h!]
+    \caption{Geographic comparison of model embedding values versus daily pollution levels}
+    \label{fig:outliers_vs_dense_cities}
+\begin{subfigure}{0.5\textwidth}
+    \centering
+    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/o3-dim1.png} 
+    \caption{O3 values of the first encoded dimension (units of compression)}
     \label{fig:outliers}
-\end{subfigure}% <- % sign must be here!
+\end{subfigure}
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.1\linewidth, height=5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/co_scatter_ae_large_cities.png}
-    \caption{Densely Populated Cities}
+    \centering
+    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/o3-first.png}
+    \caption{O3 pollution values for the first day of data set ($\mu g/m^3$)}
     \label{fig:dense_cities}
 \end{subfigure}
-
-\caption{Scatter plots of outlier and highly-populated cities over first two dimensions}
-\label{fig:outliers_vs_dense_cities}
+\begin{subfigure}{0.5\textwidth}
+    \centering
+    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/o3-dim2.png} 
+    \caption{O3 values of the second encoded dimension (units of compression)}
+    \label{fig:outliers}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+    \centering
+    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/o3-last.png}
+    \caption{O3 pollution values for the last day of data set ($\mu g/m^3$)}
+    \label{fig:dense_cities}
+\end{subfigure}
 \end{figure}
 
-\par By selecting from a list the top two hundred most highly-populated cities in the US, we found that nearly all of them were located near the center of the clustering.
-Conversely, we found that many of the outliers were smaller cities in rural locations.
+\newpage
 
@@
 
-Although the model provides insights on the most succinct features that characterize urban air pollution, more specific external data is needed to uncover these features.
+\par The proposed model provides us with many insights on the most succinct features that characterize urban air pollution in US cities.
+Based on location identifiers such as city name, state, country, and the geograpic coordinates, we are able to infer much of the explainability of the model.
+However, more specific external data is needed to fully uncover these features.
 One study \parencite{Bai2019} uses a stacked autoencoder approach and focuses on the seasonality of pm2.5 levels.
-The proposed model considers both the seasonality and temporality of pm2.5 levels and supporting meteorological data.
+The model considers both the seasonality and temporality of pm2.5 levels and supporting meteorological data.
 The research group was able to accurately predict the pm2.5 levels to the hour for a given season.
+The pollution autoencoder models we created only consider half a year of data, so we were unable to fully account for seasonal variations in the dataset.
+\par Another study focusing on pm2.5 \parencite{doreswamy2021} uses autoencoders for spatio-temporal clustering analysis.
+The group looks at Taiwan for its massive air pollution problem, coupled with the fact that the area's unique topography makes it difficult to map the distribution of pollution.
+The model uses aproximately six years of censor data from Eastern Taiwan to analyze the spatio-temporal qualities of pm2.5, as well as the socioeconomic effects of its presence.
 

### 2022-02-08 src/autoencoder.py

@@ -4,6 +4,7 @@ import itertools
 import csv
 import os
 import linreg # src/linreg.py
+import tensorflow as tf
 from tensorflow.keras.models import Model
 from tensorflow.keras.layers import Dense, Input
 from tensorflow.keras.optimizers import Adam
@@ -71,16 +72,15 @@ def ae_train_test(dims, X_train, X_test, Y_train, Y_test, component, param_grid)
         dims: Number of dimensions to train
         X_train, X_test, Y_train, Y_test: train/test data used for model
         component: Gas/particulate
-        param_grid: List of key hyperparameters to test 
+        param_grid: List of key hyperparameters to test with (found by gs)
     '''
 
     # Metrics file to write to
-    file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/ae/{component}_metrics'
+    file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/ae/{component}_metrics.csv'
     # Train Autoencoder model
     variance_list = []
     r2_list = []
-    num_of_dims = list(range(2,dims+1))
-
+    num_of_dims = list(range(1,dims+1))
     # Define hyperparams and counter
     lr = param_grid['lr'][0]
     batch = param_grid['batch'][0]
@@ -124,7 +124,7 @@ def ae_train_test(dims, X_train, X_test, Y_train, Y_test, component, param_grid)
     metrics_data.to_csv(path_or_buf=file_name, index=False)
 
 
-def ae_run(dim, X, X_train, Y_train, optimal_hyperparams, component, cities):
+def ae_run(dim, X, X_train, Y_train, optimal_hyperparams, component, label_tup):
     '''
     Run the autoencoder model by performing a cross-validated linear
     regression on the encoded data
@@ -134,7 +134,7 @@ def ae_run(dim, X, X_train, Y_train, optimal_hyperparams, component, cities):
         X: Normalized data to encode
         X_train, Y_train: Train data used in model creation
         component: Gas/particulate 
-        cities: List of cities to append to the embedding
+        label_tup: Tupple of cities, states, and countries to append to embeddings
         param_grid: The correct hyperparameters to be used. Note the dimensions that were tested in the 
             grid search. If dimension was not tested for it, use hyperparams of last tested dimension
     '''
[…]

### 2022-02-08 src/pca.py

@@ -192,24 +192,28 @@ def main():
 
     ### Input files ###
     # Open normalized data and dependent, non-normalized data
-    #dfx = pd.read_csv(f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/data/data_norm/co_data_norm.csv")
-    #dfy = pd.read_csv(f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/data/data_clean/co_data_clean.csv")
+    dfx = pd.read_csv(f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/data/data_norm/co_data_norm.csv")
+    dfy = pd.read_csv(f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/data/data_clean/co_data_clean.csv")
     # City names to append
-    #cities = dfy['city'].values
+    cities = dfy['city'].values
     
+    # Drop df labels
+    dfx.drop(['city', 'state', 'country'], axis=1, inplace=True)
     # Set x as the normalized values, y (non-normalized) as the daily average of final day
-    #X = dfx.values
-    #Y = dfy.loc[:, ['co_2021_06_06']].values
+    X = dfx.values
+    Y = dfy.loc[:, ['co_2021_06_06']].values
     # Split into train/test data
-    #X_train, X_test, Y_train, Y_test = train_test_split(X, Y, test_size=0.20, random_state=40)
+    X_train, X_test, Y_train, Y_test = train_test_split(X, Y, test_size=0.20, random_state=40)
 
     ### Function calls ###
     #interpolate(dims, component_test)
     #for component in component_names:
     #    interpolate(dims, component)
     #    pca_train_test(dims, X, Y, folds, component)
-    #pca_run(dims, X, X_train, Y_train, component_test, cities)
-    
+
+    pca_run(dims, X, X_train, Y_train, component_test, cities)
+
+    ''' 
     for component in component_names:
         # Open normalized data and dependent, non-normalized data
         dfx = pd.read_csv(f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/data/data_norm/{component}_data_norm.csv")
@@ -221,8 +225,9 @@ def main():
         # Split into train/test data
         X_train, X_test, Y_train, Y_test = train_test_split(X, Y, test_size=0.20, random_state=40)
         print(f'\n{component}')
-        variance, r2 = linreg.regression(X_train, X_test, Y_train, Y_test)
-        
[…]

### 2022-02-08 src/reverse_geocode.py

@@ -38,12 +38,13 @@ def append_labels():
     component_names = ['co', 'no', 'no2', 'o3', 'so2', 'pm2_5', 'pm10', 'nh3']
     data_labels = pd.read_csv('/home/nick/Downloads/revised_data/data_labels.csv')
     for component in component_names:
-        f_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/data_clean/{component}_data_clean.csv'
-        output_file = f'/home/nick/Downloads/revised_data/{component}_test.csv'
+        f_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/data_norm/{component}_data_norm.csv'
+        #output_file = f'/home/nick/Downloads/revised_data/{component}_test.csv'
         df = pd.read_csv(f_name)
-        df.insert(1, 'state', data_labels['state'])
-        df.insert(2, 'country', data_labels['country'])
-        df.to_csv(output_file, index=False)
+        df.insert(0, 'city', data_labels['city'])
+        #df.insert(1, 'state', data_labels['state'])
+        #df.insert(2, 'country', data_labels['country'])
+        df.to_csv(f_name, index=False)
 
 
 append_labels()

### 2022-02-08 src/spearman.py

@@ -1,0 +1,15 @@
+import pandas as pd
+import numpy as np
+from scipy.stats import spearmanr
+
+# Compare monotonicity of norm data and pca embeddings
+dfx = pd.read_csv('/home/nick/github_repos/Pollution-Autoencoders/data/data_norm/o3_data_norm.csv')
+dfy = pd.read_csv('/home/nick/github_repos/Pollution-Autoencoders/data/vec/o3_vec_190.csv')
+
+# Only compare first 10 dims with first 10 features
+dfx = dfx.iloc[:, 3:-1]
+dfy = dfy.iloc[:, 3:]
+for i in range(len(dfx.columns)):
+    rho, p = spearmanr(dfx.iloc[i], dfy.iloc[i])
+    if p < 0.05:
+        print(f'dim {i}, rho: {rho}, p: {p}')

### 2022-02-08 src/visualization.py

@@ -4,7 +4,9 @@ import matplotlib.pyplot as plt
 import seaborn as sns
 import itertools
 import pyparsing
-from mpl_toolkits import mplot3d
+import mplcursors
+from mpl_toolkits import mplot3d 
+import plotly.express as px
 
 
 def linegraph(type, dims, component_names, colors_list):
@@ -19,7 +21,6 @@ def linegraph(type, dims, component_names, colors_list):
         colors_list: List of colors to be used
     '''
     
-    num_of_comp = list(range(2,dims+1)) # TODO: delete this later
     num_of_dims = list(range(2, dims+1))
     # Plot results
     plt.figure(figsize=(12,10))
@@ -51,16 +52,6 @@ def linegraph(type, dims, component_names, colors_list):
         ax.set_xlabel('Dimension', fontsize='large')
         ax.set_ylabel('% Explained Variance', fontsize='large')
         plt.title(plt_title)
-
-        #plt.plot(num_of_comp, variance[:dims-1], label = '{}'.format(component), linestyle = '-', marker = '+', color = colors_list[i])
-        #plt.plot(num_of_comp, r2[:dims-1], linestyle = '-.', marker = 'H', color = colors_list[i])
-        #plt.rcParams.update({'font.size': 22})
-        #plt.tick_params(axis='both', which='major', labelsize=28)
-        #plt.xlabel('Dimension')
-        #plt.ylabel('% Explained Variance')
-        #plt.xlabel('')
-        #plt.ylabel('')
-        #plt.title(plt_title)
         plt.ylim([0,1])
         plt.legend()
     plt.show()
@@ -170,21 +161,21 @@ def heatmap(component):
     fig.colorbar(heat)
     plt.show()
 
+
 def correlation(X_data_file, Y_data_file, component):
     SIZE = 190
     
     # Read in normalized data
     norm_data = pd.read_csv(filepath_or_buffer=X_data_file)
    
-    X_matrix, Y_matrix = [], []
+    X_matrix = []
     norm_labels = ['dim_{}'.format(i) for i in range(1, 190+1)]
     
     
     for i in range(SIZE):
         X_matrix.append(norm_data[norm_labels[i]][:])
-    
+    print('X_matrix', X_matrix) 
     coefs = np.corrcoef(X_matrix)
-    print(coefs.shape)
 
     mask = np.zeros_like(coefs)
     mask[np.triu_indices_from(mask)] = True
[…]

### 2022-02-08 test.py

@@ -1,10 +1,0 @@
-
-x=2437
-y=875
-
-while (x!=y):
-    if (x>y):
-        x=x-y
-    elif (x<y):
-        y=y-x
-    print(f'x: {x}, y: {y}')

### 2022-06-08 paper-draft/paper.tex
Update paper draft

@@
 \usepackage{subcaption}
+
 \addbibresource{ref.bib}
@@
 \newcommand\note[1]{\textbf{\textcolor{red}{#1}}}
-\newcommand{\comment}[1]{}
 
@@
 \begin{document}
-
 % Indentation
@@
 \begin{abstract}
-Climate change is one of the most complex challenges humanity has ever faced. It is a
-multi-dimensional problem that poses economic, social, scientific, political, and 
-moral questions that must be addressed. One of the most important ways of 
-combating climate change is to research pollution mitigation strategies 
-so that policy makers and government planners are better equipped to  make 
-informed decisions. We focused on the principle cause of global warming:
-the polluting gases responsible for the rapid shift in global climate.
-We characterize cities by the amount of pollution present using an undercomplete autoencoder
-model. By representing the level of eight toxic gases present in numerous cities
-across the United States, we explore some of the more prominent features 
-that characterize urban pollution. We compare two techniques for reducing 
-the dimensionality of our input - Principle Component Analysis (PCA) and
-Autoencoders. The first two hidden dimensions represented
-a large percentage of the explained variance in the model, so we cluster
-the cities in an effort to extract any useful features. 
+
+One of the most important ways of combating climate change is to better understand the way that polluting gas levels are geographically distributed.
+We characterize cities by the amount of pollution present using an autoencoder model.
+By representing the level of eight toxic gases present in numerous cities across the Uninited States, we explore some of the more prominent features that characterize urban air pollution.
+We compare two techniques for reducing the dimensionality of our input - Principle Component Analysis (PCA) and a single-layer autoencoder model.
+We found that the autoencoder was able to explain a larger percentage of the explainable variance for nearly all of the eight gases and particulates.
+Moreover, we found that the first two hidden dimensions represented 50\% of the total explainable variance for Tropospheric Ozone ($O_3$).
+Because of this, we compared the clustering of cities along the two axes using several geographic categories, including by region and by population.
+We found that although there was significant clustering by region, there was no such clustering for the top two hundred most populous U.S.
+cities.
+
+%Climate change is one of the most complex challenges humanity has ever faced.
+It is a multi-dimensional problem that poses economic, social, scientific, political, and moral questions that must be addressed.
+One of the most important ways of combating climate change is to research pollution mitigation strategies so that policy makers and government planners are better equipped to make informed decisions.
+We focused on the principle cause of global warming: the polluting gases responsible for the rapid shift in global climate.
+We characterize cities by the amount of pollution present using an undercomplete autoencoder model.
+By representing the level of eight toxic gases present in numerous cities across the United States, we explore some of the more prominent features that characterize urban pollution.
+We compare two techniques for reducing the dimensionality of our input - Principle Component Analysis (PCA) and Autoencoders.
+The first two hidden dimensions represented a large percentage of the explained variance in the model, so we cluster the cities in an effort to extract any useful features.
+\note{Must have quantifiable results, update later!}
 \end{abstract}
@@
 \vspace{0.25cm}
+
 \section{Introduction}
 
-\par For the past several decades, air pollution has become a major concern for human health
-and the environment. The EPA sets national air quality standards for  six major air
-toxins: ground level Ozone (O$_3$), Carbon Monoxide (CO), Sulfur Dioxide (SO$_2$),
-Lead (Pb), Nitrogen Dioxide (NO$_2$), and particulate matter (PM).
+\par For the past several decades, air pollution has become a major concern for human health and the environment.
+The EPA sets national air quality standards for  six major air toxins: ground level Ozone (O$_3$), Carbon Monoxide (CO), Sulfur Dioxide (SO$_2$), Lead (Pb), Nitrogen Dioxide (NO$_2$), and particulate matter (PM).
 The sources of these emissions vary, but the two largest contributors are industrial processes and motorized vehicles.
@@
 By creating an embedding of time series pollution data for numerous cities throughout the United States, we explore the more interesting features of the model as well as apply it to tangential areas such as the pollution generated by industrial farming and animal husbandry activities.
-
 \par In this paper we use two different methods for data compression, autoencoders and principle component analysis (PCA).
@@
 Once the training process is complete, the trained models are employed to create embeddings for various encoder dimensions.
-
 \par Given the relatively short time span of our features along with the fact that pollution levels do not tend to vary drastrically day-to-day, it is evident that there is a strong correlation among features.
@@
 \section{Related Work}
-\par The idea of generating embeddings for a location is a highly potential expanse for
-research. The two major forces behind its competency are the fact that large volumes of
-data is being effectively compressed and that these embeddings, when plugged into any
-model could effectively represent any given location.  A lot of research is being done 
-on location embeddings in recent times. 
-\par For instance, location embeddings are being 
-generated based on human-mobility data \parencite{GeoEmbeddings} to understand the 
-behavioral proximity between different locations irrespective of their spatial distances.
-The sequence of locations that are frequently travelled were fed to a Skip-gram Word2vec 
-model to generate the location embeddings. 
-\par Another approach that was implemented to 
-leverage location embeddings is GPS2Vec  \parencite{GPS2Vec}.  This application consists
-of a neural network that would generate embeddings for a location, based on the semantic 
-contexts that are sourced from the user content published on multiple digital platforms.
-With the integration of these embeddings, a successful geo-tagged image classification 
-task is demonstrated. 
-\par In another effort\parencite{PlaceDeduplication}, the location embeddings derived from user content such as addresses, phone numbers, images, place
-names and all other details from multiple individual online posts, are compressed to 
-form low dimensional embeddings to solve the place duplication issue. Deduplication of 
-locations, is a paramount task for any digital platform since they provide a place graph
-to its users, which is the result of merging location data from multiple sources. This 
-merge leads to duplicity since different sources could hold different names for the same 
-place. By applying K-NN search, pair-wise duplication prediction on these generated 
-location embeddings, the application can identify the duplicates and eliminate them to 
-result in an efficient place graph.
-\par In 2018, a new approach, “DeepMove” is proposed \parencite{DeepMove}. This 
-implementation employs Trip model and OD model to learn latent representations of 
-places.  The approach is successfully applied to New York city yellow taxi dataset – 
-“includes pick-up/drop-off times, the number of passengers, pick-up/drop-off locations 
-represented by longitude and latitude, trip distance, rate types, payment type, and
-itemized fares” \parencite{DeepMove}. Similarities among locations in New York city are 
-derived effectively with “DeepMove”. To resolve the dissonance that arises from the 
-fusion of geographic and semantic features of a location, an unsupervised learning 
-implementation had been demonstrated that would derive location embeddings from human 
-trajectories \parencite{DisentangledLocationEmbeddings}. 
-\par With the ever-growing 
-footprint of digital and social media platforms in everyday life of an individual, a
-lot of research is happening to generate embeddings based on the content posted in 
-different locations. LeGo-CM is one such effort that would generate embeddings of 
-geotagged tweets \parencite{GeotaggedTweets}. The resulting embeddings could be 
-applied to understand the kind of content each location reflects. 
-\par Venue2Vec is another 
-idea that is implemented to enhance the location-aware services in various platforms 
-\parencite{Venue2Vec}. This model is trained on semantic data, temporal-spatial context 
-and sequential relations among locations to generate embeddings. This makes sure that 
-locations with similar data would fall close, when plotted in a multi-dimensional space,
-based on the input features. Hence, the model is effective in predicting the next and  
-future check-in location. 
-\par Community Flow prediction is another area of interest where
-embeddings are a major problem solver. Geo-contextual Multitask Embedding Learner (GMEL)
-, is a model that uses “Graph Attention Network” to encode geographical contexts and 
-spatial correlations to generate location embeddings \parencite{CommutingFlowPrediction}.
-These embeddings are further used to predict the community fow that serves as a vital
-factor in policy establishment and metro planning.
+\par The idea of generating embeddings for a location is a highly potential expanse for research.
+The two major forces behind its competency are the fact that large volumes of data is being effectively compressed and that these embeddings, when plugged into any model could effectively represent any given location.
+A lot of research is being done on location embeddings in recent times.
+\par For instance, location embeddings are being generated based on human-mobility data \parencite{GeoEmbeddings} to understand the behavioral proximity between different locations irrespective of their spatial distances.
+The sequence of locations that are frequently travelled were fed to a Skip-gram Word2vec model to generate the location embeddings.
+\par Another approach that was implemented to leverage location embeddings is GPS2Vec  \parencite{GPS2Vec}.
+This application consists of a neural network that would generate embeddings for a location, based on the semantic contexts that are sourced from the user content published on multiple digital platforms.
+With the integration of these embeddings, a successful geo-tagged image classification task is demonstrated.
+\par In another effort\parencite{PlaceDeduplication}, the location embeddings derived from user content such as addresses, phone numbers, images, place names and all other details from multiple individual online posts, are compressed to form low dimensional embeddings to solve the place duplication issue.
+Deduplication of locations, is a paramount task for any digital platform since they provide a place graph to its users, which is the result of merging location data from multiple sources.
+This merge leads to duplicity since different sources could hold different names for the same place.
+By applying K-NN search, pair-wise duplication prediction on these generated location embeddings, the application can identify the duplicates and eliminate them to result in an efficient place graph.
+\par In 2018, a new approach, “DeepMove” is proposed \parencite{DeepMove}.
+This implementation employs Trip model and OD model to learn latent representations of places.
+The approach is successfully applied to New York city yellow taxi dataset – “includes pick-up/drop-off times, the number of passengers, pick-up/drop-off locations represented by longitude and latitude, trip distance, rate types, payment type, and itemized fares” \parencite{DeepMove}.
+Similarities among locations in New York city are derived effectively with “DeepMove”.
+To resolve the dissonance that arises from the fusion of geographic and semantic features of a location, an unsupervised learning implementation had been demonstrated that would derive location embeddings from human trajectories \parencite{DisentangledLocationEmbeddings}.
+\par With the ever-growing footprint of digital and social media platforms in everyday life of an individual, a lot of research is happening to generate embeddings based on the content posted in different locations.
+LeGo-CM is one such effort that would generate embeddings of geotagged tweets \parencite{GeotaggedTweets}.
+The resulting embeddings could be applied to understand the kind of content each location reflects.
+\par Venue2Vec is another idea that is implemented to enhance the location-aware services in various platforms \parencite{Venue2Vec}.
+This model is trained on semantic data, temporal-spatial context and sequential relations among locations to generate embeddings.
+This makes sure that locations with similar data would fall close, when plotted in a multi-dimensional space, based on the input features.
+Hence, the model is effective in predicting the next and future check-in location.
+\par Community Flow prediction is another area of interest where embeddings are a major problem solver.
+Geo-contextual Multitask Embedding Learner (GMEL), is a model that uses “Graph Attention Network” to encode geographical contexts and spatial correlations to generate location embeddings \parencite{CommutingFlowPrediction}.
+These embeddings are further used to predict the community fow that serves as a vital factor in policy establishment and metro planning.
 
@@
 
-\par Two different methods were used to reduce the dimensionality of the data - principle component analysis (PCA) and autoencoders.
-A simple linear regression was performed on the encoded data that was generated by both the PCA and autoencoder models.
-Additionally, a linear regression on the unencoded values was performed for comparison.
-Each model's goodness of fit was evaluated using the percentage of explained variance of the three linear regressions.
-
-\par In order to explore the nature of the dataset, we calculated the Pearson correlation coefficient of the normalized timeseries data.
-The Pearson correlation coefficient allows us allows us to observe the strength of the linear correlation between features.
-
-\begin{equation*}
-    r_{XY}=\frac{cov(X,Y)}{\sigma_{X} \cdot \sigma_{Y}}
-\end{equation*}
-
-% Correlation Matrix 
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.7\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/corr_matrix.png}
-\end{center}
-    \caption{Correlation Matrix for Carbon Monoxide (daily averages for cities throughout the US)}
-\label{fig:matrix}
-\end{figure}
+\par In order to execute the proposed idea of pollution data compression, we orchestrated the data through multiple models.
+We began by applying a linear regression the uncompressed data, which gave us the requisite baseline performance scores to compare the embeddings with.
+Next, we tuned, trained, and tested eight autoencoder (AE) models and eight principle component analysis (PCA) models.
+The model metrics of the AE and PCA models are evaluated, and the resulting embeddings are applied to two further linear regressions.
+Lastly, the scores of the AE and PCA models are compared against the baseline scores to evaluate the performance, efficiency and quality of the embeddings.
 
@@
 \newpage
+
 \begin{Center}
@@
 \subsection{Data Collection and Preprocessing}
-The dataset was created using OpenWeather's Air Pollution API. The dataset includes data for 
-eight polluting gases and particulates: CO, NO, NO$_2$, O$_3$, SO$_2$, NH$_3$, PM$_{2.5}$, 
-and PM$_{10}$.
-Also included was the city name, hourly timestamp, the latitude and longitude (part of the input), and the air quality index (aqi).
-We chose not to use the aqi because it is overall less sensitive to change and is unlikely to reveal any long-term patterns over such a short time span .
-\par We looked at data for the eight pollutants over a five month time frame: January 27th, 2020 through June 6th, 2021.
+
+\par The dataset was created using OpenWeather's Air Pollution API.
+The API offers approximately six months of historical pollution data with hourly granularity.
+The hourly pollution data does not fluctuate drastically over a given day, so we opted to derive the daily average.
+Doing this made the feature set proportional to the number of cities we had data for; it also reduced the computational load without signifcantly affecting performance.
+
+\par In a preemptive effort to explore the dataset, we calculated the Pearson correlation coefficient of the normalized temporal data.
+The Pearson correlation coefficient allows us allows us to observe the strength of the linear correlation between features.
+
+\begin{equation*}
+    r_{XY}=\frac{cov(X,Y)}{\sigma_{X} \cdot \sigma_{Y}}
+\end{equation*}
+
+\par The correlation matrix of the resultant features illustrates that there are some features with strong correlation bonds, largely toward the end of the feature list.
+This reflects the fact that the average proportion of the pollutant is similar on some days.
+The short term consecutive increase in correlation demonstrates that the sources of emission follow a pattern over certain months of the year.
+Another possibility is that the pollution levels are influenced by some natural phenomenon.
+Either possibility is difficult to prove given the relatively short time frame of our data.
+However, there is evidence suggesting that temperature increases have a positive correlation with rising pollution levels, which could explain the stronger correlation present in the summer months of the dataset \parencite{jayamurugan2013}.
+
+% Correlation Matrix 
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=0.7\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/corr_matrix.png}
+\end{center}
+    \caption{Correlation Matrix for Carbon Monoxide (daily averages for cities throughout the US)}
+\label{fig:matrix}
+\end{figure}
+
+\par We looked at data for the eight pollutants over a six month time frame: January 27th, 2020 through June 6th, 2021.
 The final dataset includes just over eighteen thousand cities in the United States and other U.S.
 territories.
-We found it easier to focus on a single gas when developing the models, so we chose to concentrate on CO during the engineering process.
+The eight polluting gases and particulates that we selected were CO, NO, NO$_2$, O$_3$, SO$_2$, NH$_3$, PM$_{2.5}$, and PM$_{10}$.
+The air quality index (aqi) was also included, but we chose not to include it because it is overall less sensitive to change and is unlikely to reveal any long-term patterns over such a short time span.
+Additionally, we found it easier to focus on a single gas when developing the models, so we chose to concentrate on O$_{3}$ during the engineering process.
 The same process was then applied to the other seven gases and particulates after the development was completed.
@@
 \end{table}
-\par OpenWeather's Pollution API ordinarily returns hourly values, but we opted to use the daily average pollution levels as our feature set in order to make the number of starting dimensons more manageable.
-The daily averages were also feature engineered to be proportional to the number of features so as to avoid the curse of dimensionality \parencite{Trunk79}.
+
 \par The original data was unit normalized so that the gradient descents could converge more quickly while still preserving the original differences in range.
@@
 
+\subsection{Model Validation}
+
+\par To perform the initial linear regression on uncompressed data, we choose the the daily pollution values of one random day from the dataset while accomodating to train the model with the rest of the features.
+K-fold cross validation is applied on the model after normalizing the training data.
+This helps with evaluating the minimum, average, and maximum regression scores of the original data that was not condensed.
+These scores create a baseline that help us in comparing the scores that were acheived after applying the reduction techniques.
+
+\begin{figure}[h!]
+    \centering
+    \caption{Unencoded R2 Scores}
+    \label{fig:pipeline}
+    \includegraphics[width=0.8\linewidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/unencoded_r2.png}
+\end{figure}
+
 \subsection{Autoencoder Model}
+
 \par The autoencoder model used consists of a single encoded and decoded layer.
@@
         \hline
-        \multicolumn{5}{c}{CO Grid Search Key Dimensions 1-10}\\
+        \multicolumn{5}{c}{CO grid search for key dimensions 1-10}\\
         \hline
         dimension & $r^{2}$ & learning rate & batch & epochs  \\
+        \hline
         1 & 0.22 & 0.01 & 64 & 10 \\
@@
 \begin{center}
-    \caption{Grid Search Results for Carbon Monoxide}
     \label{tab:table2}
@@
         \hline
-        \multicolumn{5}{c}{CO Grid Search Key Dimensions 20-120}\\
+        \multicolumn{5}{c}{CO grid search for key dimensions 20-120}\\
         \hline
         dimension & $r^{2}$ & learning rate & batch & epochs  \\
+        \hline
         20 & 0.62 & 0.01 & 64 & 50\\
@@
 We prioritized the first ten dimensions because they generally represent the largest increase in the percent explained variance.
-
-\subsection{Additional Model Validation}
-In addition to the PCA and Autoencoder models, a linear regression was performed on the unencoded features.
-The resulting explainable variance scores serve as a benchmark with which we can compare each model's data loss.
-
-\begin{figure}[h!]
-    \centering
-    \caption{Unencoded R2 Scores}
-    \label{fig:pipeline}
-    \includegraphics[width=0.8\linewidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/unencoded_r2.png}
-\end{figure}
-
-\newpage
 
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/region.png} 
+    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/region.png} 
     \caption{Model characterization grouped by region}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/outside-us.png}
+    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/outside-us.png}
     \caption{Model characterization of U.S. Territories}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/california.png} 
+    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/california.png} 
     \caption{Model characterization of California}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/texas.png} 
+    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/texas.png} 
     \caption{Model characterization of Texas}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/new-york.png}
+    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/new-york.png}
     \caption{Model characterization of New York}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=0.8\linewidth, height=6.5cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/minnesota.png}
+    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/minnesota.png}
     \caption{Model characterization of Minnesota}
@@
 
+\newpage
 \par The model's characterization appears to be heavily dependent on geographic location.
@@
 However, more specific external data is needed to fully uncover these features.
-One study \parencite{Bai2019} uses a stacked autoencoder approach and focuses on the seasonality of pm2.5 levels.
+\par One study \parencite{Bai2019} uses a stacked autoencoder approach and focuses on the seasonality of pm2.5 levels.
 The model considers both the seasonality and temporality of pm2.5 levels and supporting meteorological data.
@@
 The pollution autoencoder models we created only consider half a year of data, so we were unable to fully account for seasonal variations in the dataset.
-\par Another study focusing on pm2.5 \parencite{doreswamy2021} uses autoencoders for spatio-temporal clustering analysis.
+\par Another study focusing on pm2.5 \parencite{Doreswamy2021} uses autoencoders for spatio-temporal clustering analysis.
 The group looks at Taiwan for its massive air pollution problem, coupled with the fact that the area's unique topography makes it difficult to map the distribution of pollution.
 The model uses aproximately six years of censor data from Eastern Taiwan to analyze the spatio-temporal qualities of pm2.5, as well as the socioeconomic effects of its presence.
+\par A third study \parencite{Ma2019} concentrates on recovering lost pollution mapping data with mobile sensing networks.
+The proposed framework separates the process of pollution generation and data sampling to allow the model to deal with irregular time and space intervals.
+To aid in this, a deep learning approach using autoencoders was chosen to handle the irregular sampling.
+Additionally, a convolutional long short-term memory (ConvoLSTM) structure was utilized because it allows for the capture of both temporal and spatial dependencies.
+The autoencoders were used to describe the conditional probability of complete data given partial observations, whereas the ConvLSTM was used as a sampling imitator to model the pollution generation.
 
 \section{Conclusion}
+
+\par This paper presents the advantages of using a deep learning approach to characterize urban pollution data.
+The bottleneck architecture of the autoencoder model allows the network to extract the most representative features of air pollution in US cities.
+Furthermore, the autoencoder model is shown to be more effective at representing lower dimensional latent spaces than principle component analysis.
+The model may be expanded to apply to many different types of polluting gasses, beyond the eight gasses and particulates that we have already analyzed.
+The generalized model allows us to codify the influences on urban air pollution, which can be useful for further predictive modeling and policy making.
 

### 2022-06-08 src/autoencoder.py

@@ -246,8 +246,9 @@ def main():
     
     ### Constants ###
     #components = ['co', 'no', 'no2', 'o3', 'so2', 'pm2_5', 'pm10', 'nh3']
-    components = ['co', 'no', 'no2', 'o3', 'pm2_5', 'pm10', 'nh3'] # minus so2
-    #components = 'so2'
+    # COMPLETED: co, no2, no 
+    components = ['o3', 'pm2_5', 'pm10', 'nh3'] # minus so2
+    #components = ['o3']
     # Starting dimensions
     dims = 190
     # K-fold folds

### 2022-06-08 src/linreg.py

@@ -1,7 +1,7 @@
 import numpy as np
 import pandas as pd
 from sklearn.linear_model import LinearRegression
-from sklearn.metrics import r2_score
+from sklearn.metrics import r2_score, mean_squared_error
 
 
 def regression(train_data, test_data, train_labels, test_labels):
@@ -29,6 +29,9 @@ def regression(train_data, test_data, train_labels, test_labels):
     # Variance and r2 scores for the regression model
     variance = regr.score(test_data, test_labels) 
     r2 = r2_score(test_labels, Y_pred)    
+    mse = mean_squared_error(test_labels, Y_pred)
+
+    print('Mean Squared Error: %.2f' % mse)
     print ('Variance score: %.2f' % variance)
     print ('R Square: %.2f' % r2)
 

### 2022-06-08 src/visualization.py

@@ -9,7 +9,7 @@ from mpl_toolkits import mplot3d
 import plotly.express as px
 
 
-def linegraph(type, dims, component_names, colors_list):
+def linegraph(f_name, dims, component_names, colors_list):
     ''' 
     Plot the explained variance across dimensions using the model results
 
@@ -25,17 +25,8 @@ def linegraph(type, dims, component_names, colors_list):
     # Plot results
     plt.figure(figsize=(12,10))
     for i, component in enumerate(component_names):
-        # AE or PCA
-        if type == 'ae':
-            #file_name = f'{f_name}{component}_metrics.csv'
-            file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/ae/{component}_metrics.csv'
-            plt_title = 'Autoencoder Reduced Representation of Air Pollutants'
-        elif type == 'pca':
-            file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/pca/{component}_metrics'
-            plt_title = 'PCA Reduced Representation of Air Pollutants'
-        else:
-            print('Type must be "ae" or "pca"')
-            quit()
+        file_name = f'{f_name}{component}_metrics.csv'
+        plt_title = 'Autoencoder Reduced Representation of Air Pollutants'
 
         # Read in model results
         model = pd.read_csv(filepath_or_buffer=file_name)
@@ -169,7 +160,7 @@ def correlation(X_data_file, Y_data_file, component):
     norm_data = pd.read_csv(filepath_or_buffer=X_data_file)
    
     X_matrix = []
-    norm_labels = ['dim_{}'.format(i) for i in range(1, 190+1)]
+    norm_labels = ['dim_{}'.format(i) for i in range(1, SIZE+1)]
     
     
     for i in range(SIZE):
@@ -414,19 +405,20 @@ def scattergeo(component):
 
 ### Function Calls ###
 
-COMPONENT_NAMES = ['co', 'no', 'no2', 'o3', 'so2', 'pm2_5', 'pm10', 'nh3']
+#COMPONENT_NAMES = ['co', 'no', 'no2', 'o3', 'so2', 'pm2_5', 'pm10', 'nh3']
+COMPONENT_NAMES = ['co', 'no', 'no2']
[…]

### 2023-02-19 paper-draft/paper.tex
Maintenance commit

@@
     \begin{center}
-    \includegraphics[width=10cm, height=13cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pipeline.png}
+    \includegraphics[width=10cm, height=13cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/pipeline.png}
     \end{center}
@@
 \begin{center}
-\includegraphics[width=0.7\textwidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/corr_matrix.png}
+\includegraphics[width=0.7\textwidth]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/corr_matrix.png}
 \end{center}
@@
     \label{fig:pipeline}
-    \includegraphics[width=0.8\linewidth]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/unencoded_r2.png}
+    \includegraphics[width=0.8\linewidth]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/unencoded_r2.png}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.7\linewidth, height=7cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
+    \includegraphics[width=0.7\linewidth, height=7cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
     \caption{PCA Dimensionality Reduction}
@@
     \centering
-    \includegraphics[width=0.7\linewidth, height=7cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
+    \includegraphics[width=0.7\linewidth, height=7cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/pca_all_dims.png} 
     \caption{Autoencoder Dimensionality Reduction\ \note{Placeholder for AE}}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/region.png} 
+    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/region.png} 
     \caption{Model characterization grouped by region}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/outside-us.png}
+    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/outside-us.png}
     \caption{Model characterization of U.S. Territories}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/california.png} 
+    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/california.png} 
     \caption{Model characterization of California}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/texas.png} 
+    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/texas.png} 
     \caption{Model characterization of Texas}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/new-york.png}
+    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/new-york.png}
     \caption{Model characterization of New York}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/minnesota.png}
+    \includegraphics[width=0.8\linewidth, height=6cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/minnesota.png}
     \caption{Model characterization of Minnesota}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.0\linewidth, height=10cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/population.png} 
+    \includegraphics[width=1.0\linewidth, height=10cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/population.png} 
     \caption{Model characterization grouped by region}
@@
 \begin{subfigure}{0.5\textwidth}
-    \includegraphics[width=1.0\linewidth, height=10cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/population_3and4.png}
+    \includegraphics[width=1.0\linewidth, height=10cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/population_3and4.png}
     \caption{Model characterization of U.S. Territories}
@@
     \centering
-    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/o3-dim1.png} 
+    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/o3-dim1.png} 
     \caption{O3 values of the first encoded dimension (units of compression)}
@@
     \centering
-    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/o3-first.png}
+    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/o3-first.png}
     \caption{O3 pollution values for the first day of data set ($\mu g/m^3$)}
@@
     \centering
-    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/o3-dim2.png} 
+    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/o3-dim2.png} 
     \caption{O3 values of the second encoded dimension (units of compression)}
@@
     \centering
-    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/github_repos/Pollution-Autoencoders/paper-draft/images/o3-last.png}
+    \includegraphics[width=0.8\linewidth, height=8cm]{/home/nick/git/Pollution-Autoencoders/paper-draft/images/o3-last.png}
     \caption{O3 pollution values for the last day of data set ($\mu g/m^3$)}

### 2023-02-19 src/autoencoder.py

@@ -76,7 +76,7 @@ def ae_train_test(dims, X_train, X_test, Y_train, Y_test, component, param_grid)
     '''
 
     # Metrics file to write to
-    file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/ae/{component}_metrics.csv'
+    file_name = f'/home/nick/git/Pollution-Autoencoders/data/model_metrics/ae/{component}_metrics.csv'
     # Train Autoencoder model
     variance_list = []
     r2_list = []
@@ -140,7 +140,7 @@ def ae_run(dim, X, X_train, Y_train, optimal_hyperparams, component, label_tup):
     '''
 
     # File to save model metrics to
-    vec_file = f'/home/nick/github_repos/Pollution-Autoencoders/data/vec/{component}_vec_{dim}.csv'
+    vec_file = f'/home/nick/git/Pollution-Autoencoders/data/vec/{component}_vec_{dim}.csv'
 
     # Train the model that will be used to create the embedding based on desired input dimensions
     encoded_layer, *_ = autoencoder(
@@ -192,7 +192,7 @@ def grid_search(x, y, folds, component, iter_dims, key_params):
     # Headers for grid search
     grid_list = ['fold', 'dim', 'variance', 'r2', 'lr', 'batch', 'epochs']
 
-    file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/grid_params/{component}_grid_params_1_9.csv'
+    file_name = f'/home/nick/git/Pollution-Autoencoders/data/grid_params/{component}_grid_params_1_9.csv'
     # Write header
     with open(file_name,'w+', newline='') as f:
         writer = csv.writer(f)
@@ -267,8 +267,8 @@ def main():
     for component in components:
         ### Setup ###
         # Open normalized data and dependent, non-normalized data
-        dfx = pd.read_csv(f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/data/data_norm/{component}_data_norm.csv")
-        dfy = pd.read_csv(f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/data/data_clean/{component}_data_clean.csv")
+        dfx = pd.read_csv(f"{os.environ['HOME']}/git/Pollution-Autoencoders/data/data_norm/{component}_data_norm.csv")
+        dfy = pd.read_csv(f"{os.environ['HOME']}/git/Pollution-Autoencoders/data/data_clean/{component}_data_clean.csv")
[…]

### 2023-02-19 src/distributed_gs.py

@@ -22,12 +22,12 @@ def main(*argv):
     key_params = list(itertools.product(lr,batch,epochs))
     ### Input files ###
     # Open normalized data and dependent, non-normalized data
-    dfx = pd.read_csv(f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/data/data_norm/co_data_norm.csv")
-    dfy = pd.read_csv(f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/data/data_clean/co_data_clean.csv")
+    dfx = pd.read_csv(f"{os.environ['HOME']}/git/Pollution-Autoencoders/data/data_norm/co_data_norm.csv")
+    dfy = pd.read_csv(f"{os.environ['HOME']}/git/Pollution-Autoencoders/data/data_clean/co_data_clean.csv")
     # City names to append
     cities = dfy['city'].values
     # Grid params to create model with
-    param_grid = pd.read_csv(f'/home/nick/github_repos/Pollution-Autoencoders/data/grid_params/hyperparams.csv')
+    param_grid = pd.read_csv(f'/home/nick/git/Pollution-Autoencoders/data/grid_params/hyperparams.csv')
     # Name of the component gas
     component = 'no2'
     file_name=''
@@ -41,19 +41,19 @@ def main(*argv):
     for args in sys.argv:   
         if args == "-k_1_5" or args == "--first-five":
             iter_dims = np.arange(1,6,1)
-            file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/grid_params/{component}_grid_params_1_5.csv'
+            file_name = f'/home/nick/git/Pollution-Autoencoders/data/grid_params/{component}_grid_params_1_5.csv'
         elif args == "-k_6_9" or args == "--six-to-nine":
             iter_dims = np.arange(6,10,1)
-            file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/grid_params/{component}_grid_params_6_9.csv'
+            file_name = f'/home/nick/git/Pollution-Autoencoders/data/grid_params/{component}_grid_params_6_9.csv'
         elif args == "-k_10_60" or args == "--ten-to-fifty":
             iter_dims = np.arange(10,51,10)
-            file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/grid_params/{component}_grid_params_10_50.csv'
+            file_name = f'/home/nick/git/Pollution-Autoencoders/data/grid_params/{component}_grid_params_10_50.csv'
[…]

### 2023-02-19 src/pca.py

@@ -39,7 +39,7 @@ def pca_train_test(dims, x, y, folds, component):
     '''
     
     # Metrics file
-    file_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/pca/{component}_metrics'
+    file_name = f'/home/nick/git/Pollution-Autoencoders/data/model_metrics/pca/{component}_metrics'
     # Headers 
     test_metrics_list = ['fold', 'dim', 'variance', 'r2']
     # Write header
@@ -87,10 +87,10 @@ def pca_train_test(dims, x, y, folds, component):
 def interpolate(dims, component):
     ''' Function to derive the best and worst fold for a given dimension '''
 
-    df = pd.read_csv(f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/pca/{component}_metrics')
+    df = pd.read_csv(f'/home/nick/git/Pollution-Autoencoders/data/model_metrics/pca/{component}_metrics')
     # Outputs
-    best_metrics = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/pca/best_worst/{component}_best_metrics.csv'
-    worst_metrics = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/pca/best_worst/{component}_worst_metrics.csv'
+    best_metrics = f'/home/nick/git/Pollution-Autoencoders/data/model_metrics/pca/best_worst/{component}_best_metrics.csv'
+    worst_metrics = f'/home/nick/git/Pollution-Autoencoders/data/model_metrics/pca/best_worst/{component}_worst_metrics.csv'
     # Lists to write to file
     best_list=worst_list=['dim', 'variance', 'r2']
 
@@ -160,7 +160,7 @@ def pca_run(dim, X, X_train, Y_train, component, cities):
         cities: List of cities to append to the embedding
     '''
 
-    vec_file = f'/home/nick/github_repos/Pollution-Autoencoders/data/vec/pca/{component}_vec.csv'
+    vec_file = f'/home/nick/git/Pollution-Autoencoders/data/vec/pca/{component}_vec.csv'
 
     # Train pca model
     model, encoded_train_data = pca(
@@ -192,8 +192,8 @@ def main():
 
     ### Input files ###
     # Open normalized data and dependent, non-normalized data
[…]

### 2023-02-19 src/preprocessing.py

@@ -206,7 +206,7 @@ def main():
     '''
 
     # Retrieve data for list of cities 
-    city_df = pd.read_csv(filepath_or_buffer=f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/data/other/city_lat_lon.csv")
+    city_df = pd.read_csv(filepath_or_buffer=f"{os.environ['HOME']}/git/Pollution-Autoencoders/data/other/city_lat_lon.csv")
     city_count = len(city_df)
     
     # Start and ending times
@@ -216,10 +216,10 @@ def main():
     # Component gases
     COMPONENT_NAMES = ['co', 'no', 'no2', 'o3', 'so2', 'pm2_5', 'pm10', 'nh3']
     # List of file names to write to
-    #DATA_OUT = [f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/{component}_test.csv" for component in COMPONENT_NAMES]
-    DATA_OUT = [f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/data/data_clean/{component}_data_clean.csv" for component in COMPONENT_NAMES]
+    #DATA_OUT = [f"{os.environ['HOME']}/git/Pollution-Autoencoders/{component}_test.csv" for component in COMPONENT_NAMES]
+    DATA_OUT = [f"{os.environ['HOME']}/git/Pollution-Autoencoders/data/data_clean/{component}_data_clean.csv" for component in COMPONENT_NAMES]
     # Output location of normalized data
-    NORM_OUT = [f"{os.environ['HOME']}/github_repos/Pollution-Autoencoders/{component}_norm_test.csv" for component in COMPONENT_NAMES]
+    NORM_OUT = [f"{os.environ['HOME']}/git/Pollution-Autoencoders/{component}_norm_test.csv" for component in COMPONENT_NAMES]
     # Dimensions to use
     DIMS = 190
     # Column names

### 2023-02-19 src/reverse_geocode.py

@@ -38,7 +38,7 @@ def append_labels():
     component_names = ['co', 'no', 'no2', 'o3', 'so2', 'pm2_5', 'pm10', 'nh3']
     data_labels = pd.read_csv('/home/nick/Downloads/revised_data/data_labels.csv')
     for component in component_names:
-        f_name = f'/home/nick/github_repos/Pollution-Autoencoders/data/data_norm/{component}_data_norm.csv'
+        f_name = f'/home/nick/git/Pollution-Autoencoders/data/data_norm/{component}_data_norm.csv'
         #output_file = f'/home/nick/Downloads/revised_data/{component}_test.csv'
         df = pd.read_csv(f_name)
         df.insert(0, 'city', data_labels['city'])

### 2023-02-19 src/spearman.py

@@ -3,8 +3,8 @@ import numpy as np
 from scipy.stats import spearmanr
 
 # Compare monotonicity of norm data and pca embeddings
-dfx = pd.read_csv('/home/nick/github_repos/Pollution-Autoencoders/data/data_norm/o3_data_norm.csv')
-dfy = pd.read_csv('/home/nick/github_repos/Pollution-Autoencoders/data/vec/o3_vec_190.csv')
+dfx = pd.read_csv('/home/nick/git/Pollution-Autoencoders/data/data_norm/o3_data_norm.csv')
+dfy = pd.read_csv('/home/nick/git/Pollution-Autoencoders/data/vec/o3_vec_190.csv')
 
 # Only compare first 10 dims with first 10 features
 dfx = dfx.iloc[:, 3:-1]

### 2023-02-19 src/visualization.py

@@ -66,12 +66,12 @@ def metrics_comparison(type, dims, component_names, colors_list):
     for i, component in enumerate(component_names):
         # AE or PCA
         if type == 'ae':
-            high_file = f'/home/nick/github_repos/Pollution-Autoencoders/data/modle_metrics/ae/best_worst/{component}_best_metrics.csv'
-            low_file = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/ae/best_worst/{component}_worst_metrics.csv'
+            high_file = f'/home/nick/git/Pollution-Autoencoders/data/modle_metrics/ae/best_worst/{component}_best_metrics.csv'
+            low_file = f'/home/nick/git/Pollution-Autoencoders/data/model_metrics/ae/best_worst/{component}_worst_metrics.csv'
             plt_title = 'Autoencoder Reduced Representation of Air Pollutants'
         elif type == 'pca':
-            high_file = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/pca/best_worst/{component}_best_metrics.csv'
-            low_file = f'/home/nick/github_repos/Pollution-Autoencoders/data/model_metrics/pca/best_worst/{component}_worst_metrics.csv'
+            high_file = f'/home/nick/git/Pollution-Autoencoders/data/model_metrics/pca/best_worst/{component}_best_metrics.csv'
+            low_file = f'/home/nick/git/Pollution-Autoencoders/data/model_metrics/pca/best_worst/{component}_worst_metrics.csv'
             plt_title = 'PCA Reduced Representation of Air Pollutants'
         else:
             print('Type must be "ae" or "pca"')
@@ -104,12 +104,12 @@ def metrics_comparison(type, dims, component_names, colors_list):
 
 def heatmap(component):
     # Read in component data and create an annotation list of cities
-    data = pd.read_csv('/home/nicks/github_repos/Pollution-Autoencoders/data/data_clean/{}_data_clean.csv'.format(component))
+    data = pd.read_csv('/home/nicks/git/Pollution-Autoencoders/data/data_clean/{}_data_clean.csv'.format(component))
     df = pd.DataFrame(data=data)
[…]
